### 2012-07-22 doc/thesis_proposal/thesis_proposal.tex
updated thesis proposal

%        File: thesis_proposal.tex
%
\documentclass[letterpaper,12pt]{article}
\usepackage[top=1.0in,bottom=1.0in,left=1.25in,right=1.25in]{geometry}
\usepackage{verbatim}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage[usenames]{color}
\usepackage[
naturalnames = true, 
colorlinks = true, 
linkcolor = black,
anchorcolor = black,
citecolor = black,
menucolor = black,
urlcolor = blue
]{hyperref}

%%---------------------------------------------------------------------------%%
\author{Stuart R. Slattery
\\ \href{mailto:sslattery@wisc.edu}{\texttt{sslattery@wisc.edu}}
}

\date{\today}
\title{Thesis Proposal:\\
  A Massively Parallel Stochastic Method for Nonlinear Problems}
\begin{document}
\maketitle

%%---------------------------------------------------------------------------%%
\section{Introduction}
In many fields of engineering and physics, nonlinear problems are a
primary focus of study. Recent focus on multiple physics systems adds
a new level of complexity to common nonlinear systems as solution
strategies change when they are coupled to other
problems. Furthermore, a desire for predictive simulations to enhance
the safety and performance of engineered systems creates a need for
extremely high fidelity computations to be performed for these coupled
systems as a means to capture effects not modeled by coarser
methods. In order to achieve this high fidelity, state-of-the-art
computing facilities must be leveraged in a way that is both efficient
and considerate of hardware-related issues. As scientific computing
moves towards exascale facilities with machines of $O(1,000,000)$
cores already coming online, new algorithms to solve these complex
problems must be developed to leverage this new hardware. Issues such
as resiliency to node failure and scaling to large numbers of cores
will be pertinent to robust algorithms aimed at this new
hardware. Considering these issues, this thesis proposes a novel
stochastic method to advance solution techniques for nonlinear
problems.

For some time, the particle transport community has been utilizing
Monte Carlo methods for the solution of transport problems
\cite{Lewis_1993}. The partial differential equation (PDE) community
has focused on various deterministic methods for solutions to linear
problems \cite{Saad_2003}. In between these two areas are a
not widely known small group of stochastic methods for solving sparse
linear systems \cite{Hammersley_1964, Halton_1962, Halton_1994,
  Evans_2003}. In recent years, these methods have been further
developed for transport problems in the form of Monte Carlo
Synthetic-Acceleration (MCSA) \cite{Evans_2009} but have yet to be
applied to more general sparse linear systems. Compared to other
methods in these regime, MCSA offers two attractive qualities; (1) the
linear problem operator need not be symmetric or positive-definite,
thereby reducing preconditioning complexity and (2) the stochastic
nature of the solution method provides a natural solution to the issue
of resiliency and is amenable to parallelization using modern methods
developed by the transport community \cite{Wagner_2011}. The
developement of MCSA as a general linear solver and the development of
a parallel MCSA method will be new and unique features of this work.

In addition to linear solver advancements, nonlinear solvers may also
benefit from a general and parallel MCSA scheme. In the engineering
community, nonlinear problems are often addressed by linearizing the
problem and using traditionaly iterative or direct methods
\cite{Tannehill_1997}. In the mathematics community, various Newton
methods have been popular \cite{Kelley_1995}. Recently, Jacobian-Free
Newton-Krylov (JFNK) schemes have been utilized in multiple physics
architectures and advanced single physics codes \cite{Knoll_2004,
  Gaston_2009}. The benefits of JFNK schemes are that the Jacobian is
never formed, simplifying the implementation, and a Krylov solver is
leveraged (typically GMRES or Conjugate Gradient) providing excellent
convergence properties for well-conditioned systems. However, there
are two potential drawbacks to these methods for high fidelity
predictive simulations: (1) the Jacobian is approximated by a
first-order differencing method on the order of machine precision that
can cause large errors for fine-grained simulation and (2) for systems
that are not symmetric positive-definite (which will be the case for
most multiphysics systems and certainly for most preconditioned
systems) the Krylov subspace generated by the GMRES solver my become
prohibitively large. To address these issues, this thesis proposes a
new and novel method for nonlinear systems based on the MCSA method.

The Forward-Automated Newton-MCSA (FANM) method is proposed as new
nonlinear solution method. The key features of FANM are: full Jacobian
generation using modern Forward Automated Differencing (FAD) methods
\cite{Bartlett_2006}, and MCSA as the inner linear solver. This method
has several attractive properties. First, the first-order
approximation to the Jacobian used in JFNK type methods is eliminated
by generating the Jacobian explicitly with the model equations through
FAD. Second, the Jacobian need not be explicitly formed by the user
but is instead automated through FAD; this eleminates the complexity
of hand-coding derivatives and has also been demonstrated to be more
efficient computationally than hand-coded derivatives
\cite{Bartlett_2006}. Third, unlike GMRES, MCSA does not build a
subspace during iterations. Although the Jacobian must be explicitly
formed to use MCSA, for problems that take more than a few GMRES
iterations to converge the size of the Krylov subspace will grow
beyond that of the Jacobian. Finally, using MCSA for the linear solve
provides its benefits for preconditioning, potential resiliency, and
parallelism.

%%---------------------------------------------------------------------------%%
\section{Deliverables}

This work will deliver the following items:

\begin{itemize}
\item Develop a massively parallel strategy for the MCSA method for
  sparse linear systems.
\item Implement a massively parallel strategy for the MCSA method for
  sparse linear systems.
\item Fully assess the performance of parallel MCSA on leadership class
  computing systems in order to determine its feasibility using
  benchmark problems.
\item Compare the performance of MCSA to other iterative methods for
  sparse linear systems including GMRES.
\item Develop the FANM method for nonlinear systems.
\item Implement the FANM method for nonlinear systems.
\item Fully assess the performance of the FANM method on leadership
  class computing systems using model problems.
\item Compare the performance of MCSA to other nonlinear methods
  including JFNK.
\end{itemize}

This work will not address the resiliency of these algorithms.

%%---------------------------------------------------------------------------%%
\section{Dissertation Outline}

Based on the above work, the dissertation can be outlined in five
distinct parts:

\begin{enumerate}
\item Iterative methods for sparse linear systems (fixed-point,
  subspace, and stochastic methods).
\item MCSA and MSOD (development and implementation).
\item Iterative methods for nonlinear systems (Newton methods).
\item FANM (developement and implementation).
\item Application to model problems (conduction, Navier-Stokes,
  conjugate heat transfer).
\end{enumerate}

%%---------------------------------------------------------------------------%%
\section{Committee}

\begin{itemize}
\item Paul Wilson (advisor): Associate Professor, Department of
  Engineering Physics.
\item Greg Moses: Professor, Department of Engineering Physics.
\item Carl Sovinec: Profssor, Department of Engineering Physics.
\item Chris Rutland: Professor, Deptartment of Mechanical Engineering.
\item Tom Evans: Senior R\&D Staff, Reactor and Nuclear Systems
  Division, Oak Ridge National Laboratory.

\end{itemize}

Potential alternates include Robert Witt and Leslie Smith.

%%---------------------------------------------------------------------------%%
\section{Timeline}

%%---------------------------------------------------------------------------%%
\pagebreak
\bibliographystyle{ieeetr}
\bibliography{references}
\end{document}