### 2012-08-21 doc/prelim/document/chapters/fanm/fanm.tex
adding parallel mcsa and fanm chapters

\chapter{The FANM Method for Nonlinear Problems}
\label{ch:fanm}