### 2012-08-21 doc/prelim/document/chapters/parallel_mcsa/parallel_mcsa.tex

\chapter{Parallelization of the MCSA Algorithm}
\label{ch:parallel_mcsa}