### 2012-08-21 doc/prelim/document/chapters/chimera_codebase/chimera_codebase.tex
updating references

\chapter{The Chimera Code Base}
\label{ch:chimera_codebase}