### 2012-08-21 doc/prelim/document/chapters/fanm/fanm.tex
adding intro and research proposal chapter

\chapter{The FANM Method for Nonlinear Problems}
\label{ch:fanm}

\section{Automatic Differentiation}
\label{sec:automatic_differentiation}

\section{Residual and Jacobian Generation}
\label{sec:fanm_generation}

\section{Algorithm Outline}
\label{sec:algorithm_outline}