### 2012-08-21 doc/prelim/document/chapters/parallel_mcsa/parallel_mcsa.tex

\chapter{Parallelization of the MCSA Algorithm}
\label{ch:parallel_mcsa}

\section{Multiple-Set Overlapping-Domain Decomposition}
\label{sec:msod}

\section{Algorithm Outline}
\label{sec:algorithm_outline}