### 2012-08-22 doc/prelim/document/chapters/chimera_codebase/chimera_codebase.tex
Finished initial prelim outline

\chapter{The Chimera Code Base}
\label{ch:chimera_codebase}

The Chimera code base will be the primary research tool for answering
the questions posed in the previous chapters of the document. We may
not need a chapter for this, but it will likely be a good idea
somewhere in this document to outline the design of the software and
the instrumentation that it will provide.