### 2012-08-24 doc/me699/proposal/me699_proposal.tex
updates to me699 proposal

%        File: me699_proposal.tex
%
\documentclass[letterpaper,12pt]{article}
\usepackage[top=1.0in,bottom=1.0in,left=1.25in,right=1.25in]{geometry}
\usepackage{verbatim}
\usepackage{amsthm} 
\usepackage{tmadd,tmath}
\usepackage[mathcal]{euscript}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage[usenames]{color}
\usepackage[
naturalnames = true, 
colorlinks = true, 
linkcolor = black,
anchorcolor = black,
citecolor = black,
menucolor = black,
urlcolor = blue
]{hyperref}

%%---------------------------------------------------------------------------%%
\author{Stuart R. Slattery
\\ \href{mailto:sslattery@wisc.edu}{\texttt{sslattery@wisc.edu}}
}

\date{\today} \title{ME699 Proposal: Fully Implicit Solutions to the
  Steady and Incompressible Navier-Stokes Equations}
\begin{document}
\maketitle

%%---------------------------------------------------------------------------%%
\section{Proposal}
Fully implicit formulations of the Navier-Stokes equations are
becoming increasingly popular due to their robustness over
conventional segregated methods that arises from a fully consistent
problem formulation. Solutions with inexact Newton methods have
proliferated due to advancements in both hardware and numerical
methods for solving sparse linear sytems and preconditioning
\cite{shadid_1997,mchugh_1994,gropp_2001,knoll_2004}. To study the
details of how fully implicit solutions are implemented in practice,
and independent study is posed to develop a piece of software aimed at
generating fully implicit solutions to the steady and incompressible
Navier-Stokes equations.

\subsection{Model Problem}
We choose the steady incompressible Navier-Stokes equations as model
problems in residual form to solve for the state vector
$\{\ve{u},P,T\}$ where $\ve{u}$ is the fluid velocity, $P$ is the
fluid pressure, and $T$ is the fluid temperature. As formulated in
\cite{shadid_1997}, we consider vector forms of momentum transport:

\begin{equation}
  \rho \ve{u} \cdot \nabla \ve{u} - \nabla \cdot \ve{T} - \rho \ve{g}
  = \ve{0} \:,
  \label{eq:momentum_transport}
\end{equation}

continuity:

\begin{equation}
  \nabla \cdot \ve{u} = \ve{0} \:,
  \label{eq:continuity}
\end{equation}

and energy transport with the effects of viscous dissipation neglected
for low Mach number flow:

\begin{equation}
  \rho C_p \ve{u} \cdot \nabla T + \nabla \ve{q} = \ve{0} \:,
  \label{eq:energy_transport}
\end{equation}

where $\rho$ is the fluid density, $\ve{g}$ the gravity vector and
$C_p$ the specific heat of the fluid at constant pressure. To close
the system we define the stress tensor:

\begin{equation}
  \ve{T} = -P \ve{I} + \mu [ \nabla \ve{u} \nabla \ve{u}^{T} ] \:,
  \label{eq:stress_tensor}
\end{equation}

and the heat flux:

\begin{equation}
  \ve{q} = -\kappa \nabla T
  \label{eq:heat_flux}
\end{equation}

where $\ve{I}$ is the identity matrix, $\mu$ is the dynamic
viscosityof the fluid, and $\kappa$ is the thermal conductivity of the
fluid. These fluid properties will be assumed constant and boundary
conditions will be chosen as a combination of Neumann and Dirichlet
conditions depending on the problem chosen.

\subsection{Discretization}
Reference \cite{shadid_1997} uses a pressure stabilized
Petrov-Galerkin finite element formulation of these equations
\cite{hughes_1986}. Depending on the software infrastructure
developed, these may or may not be as easy to formulate as the finite
volume discretizations in \cite{mchugh_1994} due to the availabilty of
finite element discretization libraries, finite element assembly
engines, and unstructured mesh databases available in the Trilinos
scientific computing libraries \cite{trilinos_2005}. It will likely be
easiest to start with the finite volume approach.

\subsection{Solution Method}
We will choose an inexact Newton method for the nonlinear scheme solve
these equations \cite{dembo_1982}. Of particular interest will be
investigating how the forcing term affects convergence of the
nonlinear resdiual \cite{eisenstat_1996}. In addition, if convergence
issues are noted, globalization methods may be investigated to improve
the Newton correction \cite{pawlowski_2006}. The linear solver for the
Newton method will be GMRES as its convergence properties for these
types of systems are well documented \cite{saad_1986}. We will seek to
solve these equations using parallel methods provided by well
established framework libraries provided by the Trilinos suite of
scientific computing libraries \cite{trilinos_2005}.

\subsection{Benchmark Problems}
To verify the implementation, three widely reference two-dimensional
benchmark problems will be assessed using the resulting software. The
natural convection cavity problem \cite{davis_1983}, the driven cavity
problem \cite{ghia_1982}, and flow over a backwards step
\cite{gartling_1990}.

%%---------------------------------------------------------------------------%%
\pagebreak
\bibliographystyle{ieeetr}
\bibliography{references}
\end{document}