### 2012-09-13 doc/prelim/document/chapters/chimera_codebase/chimera_codebase.tex
updates to prelim, projection methods section is complete

\chapter{The Chimera Code Base}
\label{ch:chimera_codebase}

The Chimera code base will be the primary research tool for answering
the questions posed in the previous chapters of the document. We may
not need a chapter for this, but it will likely be a good idea
somewhere in this document to outline the design of the software and
the instrumentation that it will provide.

\section{Finite Element Assembly}
\label{ch:fem_assembly}

\section{Graph-Based Multiphysics Formulation}
\label{ch:multiphysics_graph}

\section{Software Engineering}
\label{ch:software_engineering}