### 2012-09-13 doc/prelim/document/chapters/fanm/fanm.tex

\chapter{The FANM Method for Nonlinear Problems}
\label{ch:fanm}

In this chapter we will devise the FANM method. Based on the
background in the previous section, we first need to motivate the
development of this work by pointing out some of the issues with
conventional methods. This includes JFNK coarseness and either
prohibitively large Krylov subspaces or a loss of orthogonality
information in the subspace due to restarts. We can than point out
some of the attractive qualities of the FANM method. Much of this will
rely on advanced concepts including automatic differentiation and
on-the-fly residual and Jacobian generation from nonlinear function
evaluations. In addition, we will want to show that the Newton method
will converge using MCSA as the linear solver. I suspect I will have
some preliminary results for a serial implementation here as I already
have a serial MCSA implemented. Finally, we will want to design the
model problems that we will test FANM with. These should be problems
that have already been worked up in the literature with benchmark-type
solutions.

\section{Automatic Differentiation}
\label{sec:automatic_differentiation}

\section{Residual and Jacobian Generation}
\label{sec:fanm_generation}

\section{Jacobian Storage vs. Subspace Storage and Restarts}
\label{sec:fanm_storage}

\section{Algorithm Outline}
\label{sec:fanm_algorithm_outline}