### 2012-09-22 doc/prelim/document/chapters/nonlinear_problem/nonlinear_problem.tex
working on the nonlinear chapter

\chapter{Stochastic Solution Methods for Nonlinear Problems}
\label{ch:nonlinear_problem}
Nonlinear equation sets are a hallmark of multiphysics
problems. Systems of partial differntial equations such as those that
describe fluid flow or more general transport processes when
discretized by conventional methods yield discrete sets of equations
with nonlinearities present in the variables. Traditionally, such
systems have been solved by linearizing them such that the
nonlinearities in the variable are eliminated and more traditional
linear methods can be used for solutions. Often characterized as
segregated methods where physics operators are split and their action
on the system approximated in steps, such methods lack consistency and
accuracy in resolving the nonlinear component of the solution. In the
last 30 years, fully consistent nonlinear methods based on Newton's
method have become more popular and many advances have been made in
the computational physics field to employ these methods. Currently,
Krylov methods are used with Newton's method to form a robust solution
sequence for these types of problems. However, given our goals of
advancing stochastic solver technology for linear problems, we aim to
apply them to these fully consistent schemes in order to gauge if some
improvements can made. We therefore devise a new method using MCSA in
conjuction with new methods, accomodating the fact that MCSA requires
an explicitly formed linear operator in its implementation. In
addition, we will review parallel strategies for a Newton-MCSA method.

\section{Preliminaries}
\label{sec:nonlinear_preliminaries}
We formulate the \textit{nonlinear problem} as follows
\citep{knoll_jacobian-free_2004}:
\begin{equation}
  \ve{F}(\ve{u}) = \ve{0}\:,
  \label{eq:nonlinear_problem}
\end{equation}
where $\ve{u}$ is the solution vector and $\ve{F}(\ve{u})$ is the
function of nonlinear residuals. We write the nonlinear system in this
form so that when an exact solution for $\ve{u}$ is achieved, all
residuals evaluate to zero. \textit{Newton's method} is a root finding
algorithm and therefore we can use it to solve
Eq~(\ref{eq:nonlinear_problem}) if we interpret the exact solution
$\ve{u}$ to be the roots of $\ve{F}(\ve{u})$. Newton's method is also
an iterative scheme, and we can generate this scheme by generating the
Taylor expansion of the $k+1$ iterate of the nonlinear residuals about
the $k$ iterate:
\begin{equation}
  \ve{F}(\ve{u}^{k+1}) = \ve{F}(\ve{u}^{k}) +
  \ve{F}'(\ve{u}^{k})(\ve{u}^{k+1}-\ve{u}^{k}) +
  \frac{\ve{F}''(\ve{u}^{k})}{2}(\ve{u}^{k+1}-\ve{u}^{k})^2 + \cdots
  \:.
  \label{eq:newton_derivation_1}
\end{equation}
If we ignore the nonlinear terms in the expansion and assert that the
$k+1$ iterate $\ve{u}^{k+1}$ is that exact solution such that
$\ve{F}(\ve{u}^{k+1}) = \ve{0}$, then we are left with the following:
\begin{equation}
  -\ve{F}(\ve{u}^{k}) =
  \ve{F}'(\ve{u}^{k})(\ve{u}^{k+1}-\ve{u}^{k})\:.
  \label{eq:newton_derivation_2}
\end{equation}
We note two things in Eq~(\ref{eq:newton_derivation_2}). The first is
that $\ve{F}'(\ve{u}^{k})$ is in fact the \textit{Jacobian},
$\ve{J}(\ve{u})$, of the set of nonlinear residuals and is defined
element-wise as:
\begin{equation}
  J_{ij} = \frac{\partial F_i(\ve{u})}{\partial u_j}\:.
  \label{eq:jacobian_def}
\end{equation}
Second, we note that $(\ve{u}^{k+1}-\ve{u}^{k})$ is simply the
\textit{Newton correction} at the $k$ iterate, $\delta \ve{u}^k$. We
can then rerarrange Eq~(\ref{eq:newton_derivation_2}) to define the
Newton iteration scheme for nonlinear problems:
\begin{subequations}
  \begin{gather}
    \ve{J}(\ve{u}) \delta \ve{u}^k = -\ve{F}(\ve{u}^{k})\\
    \ve{u}^{k+1} = \ve{u}^k + \delta \ve{u}^k\:.
  \end{gather}
  \label{eq:newton_iteration}
\end{subequations}
We then have two distinct steps to perform: the solution of a linear
system to compute the Newton correction where the Jacobian matrix of
the nonlinear equation set is the linear operator, and the application
of the correction to the previous iterate's solution to arrive at the
next iterate's solution. In the asymptotic lmiit, the iterations of
Newton's method will converge the nonlinear residual quadratically
\citep{kelley_iterative_1995}. Convergence criteria is set for
stopping the iteration sequence based on the nonlinear
residual. Commonly the following criteria is used:
\begin{equation}
  ||\ve{F}(\ve{u}^{k})|| < \epsilon ||\ve{F}(\ve{u}^{0})||\:,
  \label{eq:newton_stopping_criteria}
\end{equation}
where $\epsilon$ is a user defined tolerance parameter. Newton's
method is \textit{consistent} in that all components of the nonlinear
functions that describe the physics we are modeling are updated
simulataneously in the iteration sequence with respect to one
another. This is in comparison to \textit{inconsistent} strategies,
such as a pressure correction strategy for solving the Navier-Stokes
equations \citep{pletcher_computational_1997}, where the components of
$\ve{u}$ are updated in a staggered fashion depending on the
particular equations that they are associated with.

\section{Inexact Newton Methods}
\label{sec:newton_methods}
Inexact Newton methods arise when the Jacobian operator is not exactly
inverted, resulting in an inexact Newton correction as initially
described by Dembo and others \citep{dembo_inexact_1982}. For common
sparse nonlinear systems, which in turn yield a sparse Jacobian
matrix, this situation occurs when conventional iterative methods are
applied.

\subsection{Newton-Krylov Methods}
\label{subsec:newton_krylov_methods}
A form of inexact Newton methods, \textit{Newton-Krylov methods} are
nonlinear iterative methods that use a Krylov subspace method as the
linear solver for generating the Newton correction. As we investigated
in Chapter~\ref{chap:linear_problem}, Krylov methods are robust and
enjoy efficient parallel implementations. Furthermore, their lack of
explicit dependence on the operator make them easier to implement than
other methods.

In most cases, the Jacobian is difficult to form from the difference
equations. For simple nonlinear cases such as the Navier Stokes
equations, the derivatives can be computed and coded, but due to the
complexity of those derivatives and the resulting difference equations
this task can be tedious, error prone, and must be repeated for every
equation set. By recognizing that Krylov methods only need the action
of the operator on the vector instead of the operator itself, the
Jacobian can instead be approximated through various numerical methods
including a difference-based Jacobian-free formulation. Jacobian-Free
methods, and in particular \textit{Jacobian-Free Newton-Krylov} (JFNK)
methods \citep{knoll_jacobian-free_2004}, rely on forming the action
of the Jacobian on a vector as required by the Krylov solver through a
forward difference scheme. In this case, the action of the Jacobian on
some vector $\ve{v}$ is given as:
\begin{equation}
  \ve{J}(\ve{u})\ve{v} = \frac{\ve{F}(\ve{u} + \epsilon \ve{v}) -
    \ve{F}(\ve{u})}{\epsilon}\:,
  \label{eq:jacobian_free_product}
\end{equation}
where $\epsilon$ is a small number typically on the order of machine
precision. Kelley \citep{kelley_iterative_1995} points out a potential
downfall of this formulation in that if the discretization error in
$\ve{F}(\ve{u})$ is on the order of the perturbation parameter
$\epsilon$, then the finite difference error from
Eq~(\ref{eq:jacobian_free_product}) pollutes the solution.

\section{The FANM Method}
\label{sec:fanm}
In this section we will devise the FANM method. Based on the
background in the previous section, we first need to motivate the
development of this work by pointing out some of the issues with
conventional methods. This includes JFNK coarseness and either
prohibitively large Krylov subspaces or a loss of orthogonality
information in the subspace due to restarts. We can than point out
some of the attractive qualities of the FANM method. Much of this will
rely on advanced concepts including automatic differentiation and
on-the-fly residual and Jacobian generation from nonlinear function
evaluations. In addition, we will want to show that the Newton method
will converge using MCSA as the linear solver. I suspect I will have
some preliminary results for a serial implementation here as I already
have a serial MCSA implemented. Finally, we will want to design the
model problems that we will test FANM with. These should be problems
that have already been worked up in the literature with benchmark-type
solutions. Parallelism in the context of the linear solvers chapter
should also be discussed.

\subsubsection{Automatic Differentiation}
\label{subsubsec:automatic_differentiation}

\subsection{Residual and Jacobian Generation}
\label{subsec:fanm_generation}

\subsection{Jacobian Storage vs. Subspace Storage and Restarts}
\label{subsec:fanm_storage}

\subsection{Parallel FANM Implementation}
\label{subsec:parallel_fanm}