### 2012-09-24 doc/prelim/document/chapters/nonlinear_problem/nonlinear_problem.tex
nonlinear updates

\chapter{Stochastic Solution Methods for Nonlinear Problems}
\label{ch:nonlinear_problem}
Nonlinear equation sets are a common occurence in multiphysics
problems. Systems of partial differntial equations such as those that
describe fluid flow or more general transport processes when
discretized by conventional methods yield discrete sets of equations
with nonlinearities present in the variables. Traditionally, such
systems have been solved by linearizing them such that the
nonlinearities in the variable are eliminated and more traditional
linear methods can be used for solutions. Often characterized as
segregated methods where physics operators are split and their action
on the system approximated in steps, such methods lack consistency and
accuracy in resolving the nonlinear component of the solution. In the
last 30 years, fully consistent nonlinear methods based on Newton's
method have become more popular and many advances have been made in
the computational physics field to employ these methods.

Currently, Krylov methods are used with Newton's method to form a
robust solution sequence for these types of problems. However, given
our goals of advancing stochastic solver technology for linear
problems, we aim to apply them to these fully consistent schemes in
order to gauge if some improvements can be made. We therefore devise a
new method using MCSA in conjuction with new methods, accomodating the
fact that MCSA requires an explicitly formed linear operator in its
implementation as compared to Krylov methods. In addition to handling
this dependency, we will review parallel strategies for a nonlinear
MCSA method.

\section{Preliminaries}
\label{sec:nonlinear_preliminaries}
We formulate the \textit{nonlinear problem} as follows
\citep{knoll_jacobian-free_2004}:
\begin{equation}
  \ve{F}(\ve{u}) = \ve{0}\:,
  \label{eq:nonlinear_problem}
\end{equation}
where $\ve{u} \in \mathbb{R}^n$ is the solution vector and
$\ve{F}:\mathbb{R}^N \rightarrow \mathbb{R}^N$ is the function of
nonlinear residuals. We write the nonlinear system in this form so
that when an exact solution for $\ve{u}$ is achieved, all residuals
evaluate to zero. \textit{Newton's method} is a root finding algorithm
and therefore we can use it to solve Eq~(\ref{eq:nonlinear_problem})
if we interpret the exact solution $\ve{u}$ to be the roots of
$\ve{F}(\ve{u})$. Newton's method is also an iterative scheme, and we
can generate this procedure by building the Taylor expansion of the
$k+1$ iterate of the nonlinear residuals about the previous $k$
iterate:
\begin{equation}
  \ve{F}(\ve{u}^{k+1}) = \ve{F}(\ve{u}^{k}) +
  \ve{F}'(\ve{u}^{k})(\ve{u}^{k+1}-\ve{u}^{k}) +
  \frac{\ve{F}''(\ve{u}^{k})}{2}(\ve{u}^{k+1}-\ve{u}^{k})^2 + \cdots
  \:.
  \label{eq:newton_derivation_1}
\end{equation}
If we ignore the nonlinear terms in the expansion and assert that at
the $k+1$ iterate $\ve{u}^{k+1}$ is that exact solution such that
$\ve{F}(\ve{u}^{k+1}) = \ve{0}$, then we are left with the following
equality:
\begin{equation}
  -\ve{F}(\ve{u}^{k}) =
  \ve{F}'(\ve{u}^{k})(\ve{u}^{k+1}-\ve{u}^{k})\:.
  \label{eq:newton_derivation_2}
\end{equation}
We note two things of importance in
Eq~(\ref{eq:newton_derivation_2}). The first is that
$\ve{F}'(\ve{u}^{k})$ is in fact the \textit{Jacobian},
$\ve{J}(\ve{u})$, of the set of nonlinear residuals and is defined
element-wise as:
\begin{equation}
  J_{ij} = \frac{\partial F_i(\ve{u})}{\partial u_j}\:.
  \label{eq:jacobian_def}
\end{equation}
Second, we note that $(\ve{u}^{k+1}-\ve{u}^{k})$ is simply the
solution update from the $k$ iterate to the $k+1$ iterate. We will
define this update as the \textit{Newton correction} at the $k$
iterate, $\delta \ve{u}^k$. We can then rerarrange
Eq~(\ref{eq:newton_derivation_2}) to define the Newton iteration
scheme for nonlinear problems:
\begin{subequations}
  \begin{gather}
    \ve{J}(\ve{u}) \delta \ve{u}^k = -\ve{F}(\ve{u}^{k})\\
    \ve{u}^{k+1} = \ve{u}^k + \delta \ve{u}^k\:.
  \end{gather}
  \label{eq:newton_iteration}
\end{subequations}
We then have three distinct steps to perform: evaluation of the
nonlinear residuals using the solution at the $k$ iterate, the
solution of a linear system to compute the Newton correction where the
Jacobian matrix of the nonlinear equation set is the linear operator,
and the application of the correction to the previous iterate's
solution to arrive at the next iterate's solution. In the asymptotic
limit, the iterations of Newton's method will converge the nonlinear
residual quadratically \citep{kelley_iterative_1995}. Convergence
criteria is set for stopping the iteration sequence based on the
nonlinear residual. Commonly, the following criteria is used:
\begin{equation}
  ||\ve{F}(\ve{u}^{k})|| < \epsilon ||\ve{F}(\ve{u}^{0})||\:,
  \label{eq:newton_stopping_criteria}
\end{equation}
where $\epsilon$ is a user defined tolerance parameter. Newton's
method is \textit{consistent} in that all components of the nonlinear
functions that describe the physics we are modeling are updated
simulataneously in the iteration sequence with respect to one
another. This is in comparison to \textit{inconsistent} strategies,
such as a pressure correction strategy for solving the Navier-Stokes
equations \citep{pletcher_computational_1997}, where the components of
$\ve{u}$ are updated in a staggered fashion depending on the
particular equations that they are associated with.

\section{Inexact Newton Methods}
\label{sec:newton_methods}
Inexact Newton methods arise when the Jacobian operator is not exactly
inverted, resulting in an inexact Newton correction as initially
described by Dembo and others \citep{dembo_inexact_1982}. For common
sparse nonlinear systems, which in turn yield a sparse Jacobian
matrix, this situation occurs when conventional iterative methods are
applied. In their definition, Dembo formulated inexact methods such
that they are independent of the linear method used to solve for the
Newton correction and therefore are amenable to use with any linear
solver. Furthermore, they bind the convergence of the outer nonlinear
iteration to the inner linear iteration such that:
\begin{equation}
  ||\ve{J}(\ve{u}^k)\delta \ve{u}^k + \ve{F}(\ve{u}^k)|| \leq \eta^k
  ||\ve{F}(\ve{u}^k)||\:,
  \label{eq:inexact_newton_forcing}
\end{equation}
where $\eta^k \in [0,1)$ is defined as the \textit{forcing term} at
  the $k$ iterate. Eq~(\ref{eq:inexact_newton_forcing}) then states
  that the residual generated by the linear solver is bound by the
  nonlinear residual and how tightly it is bound is defined by the
  forcing term. This is useful in that we can vary how tightly coupled
  the convergence of the linear iterations used to generate the Newton
  correction is to the nonlinear iteration by relaxing or tightening
  the convergence properties on the linear iterative method. As a
  result, strategies for determining the forcing term can vary
  depending on the problem type and can greatly affect the convergence
  of the method or even prohibit convergence
  \citep{eisenstat_choosing_1996}. In addition, globalization methods
  may be used to modify the Newton correction in a more desireable
  direction such that convergence properties can be improved when the
  initial guess for $\ve{u}$ is poor
  \citep{pawlowski_globalization_2006}.

\subsection{Newton-Krylov Methods}
\label{subsec:newton_krylov_methods}
A form of inexact Newton methods, \textit{Newton-Krylov methods} are
nonlinear iterative methods that leverage a Krylov subspace method as
the linear solver for generating the Newton correction
\citep{kelley_iterative_1995}. As we investigated in
Chapter~\ref{ch:linear_problem}, Krylov methods are robust and enjoy
efficient parallel implementations on modern
architectures. Furthermore, their lack of explicit dependence on the
operator make them easier to implement than other
methods. Additionally, although many iterations can become memory
intensive due to the need to store the Krylov subspace for the
orthogonalization procedure, at each nonlinear iteration this cost is
reset as the Jacbobian matrix will change due to its dependence on the
solution vector. This means that for every nonlinear iteration, a
completely new linear system is formed for generating the Newton
correction and we can modify the Krylov solver parameters accordingly
to accomodate this. In most nonlinear problems, the Jacobian operator
is generally nonsymmetric and therefore either Krylov methods with
long recurrence relations that can handle nonsymmetric systems must be
considered or Newton correction system must be preconditioned such
that the operator is symmetric and short recurrence relations method
can be potentially be used.

With many Krylov methods available, which to use with the Newton
method is dependent on many factors including convergence rates and
memory usage. Several studies have been performed to investigate this
\citep{mchugh_inexact_1993,knoll_newton-krylov_1995}. In their
numerical studies in 1995, Knoll and McHugh used the set of highly
nonlinear and stiff convection-diffusion-reaction equations to solve a
set of tokamak plasma problems with the goal of measuring solver
performance with Newton's method. They note several tradeoffs in using
Krylov methods the a Newton solver. The first is that the optimization
condition that results from the constraints (e.g. the minimization of
the GMRES residual over the Krylov space) can be relaxed by
restricting the size of the subspace such that only a fixed number of
subspace vectors may be maintained, thus reducing memory
requirements. We can also relax the optimization condition instead
restarting the recurrence relation with a new set of vectors once a
certain number of vectors have been generated, the optimization
condition is maintain over that particular set of vectors, however,
Knoll and McHugh note that this ultimately slows the convergence rate
as compared to keeping all vectors as the new set of vectors is not
necessarily orthogonal to the previous set, and therfore not optimal
over the entire iteration procedure. The orthogonality condition can
be relaxed by using a recurrence relation that does not generate a
strictly orthonormal basis for the Krylov subspace such as the Lanzcos
biorthogonalization procedure, resulting in memory savings due to the
shorter Lanzcos recurrence relation.

As a comparison, Knoll and McHugh chose an Arnoldi-based GMRES with a
fixed vector basis approximately the size of the number of iterations
required to converge as the long recurrence relation solver and
conjugate gradients squared (CGS), bi-orthogonalized conjugate
gradient stabilized (Bi-CGSTAB), and transpose-free quasiminimal
residual (TFQMR) methdos as Lanzcos-based short recurrence relation
solvers. All solvers were used to compute the right-preconditioned
Newton correction system. For standard implementations of Newton's
method where the Jacobian operator was explicitly formed using
difference equations, all methods exhibited roughly equivalent
iteration count performance for both the inner linear iterations and
the outer nonlinear iterations in terms of iterations required to
converge with Bi-CGSTAB typically performing the best for
implementations where the Jacobian was explicitly formed and GMRES
perfoming best for matrix-free implementations. However, upon
investigating the convergence of the inner iterations, it was observed
that the GMRES solver was significantly more robust, always generating
a monotonically decreasing residual as compared to the Lanzcos-based
methods which had the tendency to oscillate. Based on these results,
in all of their future work Knoll and McHugh tended to use GMRES as
the Krylov solver \citep{knoll_jacobian-free_2004}.

\subsubsection{Jacobian-Free Approximation}
\label{subsubsec:jacobian_free_approximation}
In most cases, the Jacobian is difficult to form from the difference
equations and costly to evaluate. For simple nonlinear cases such as
the Navier-Stokes equations, the derivatives can be computed and
coded, but due to the complexity of those derivatives and the
resulting difference equations this task can be tedious, error prone,
and must be repeated for every equation set. Further, in their 1995
work, Knoll and McHugh also noted that a dominating part of their
computation time was the evaluation of the difference equations for
building the Jacobian \citep{knoll_newton-krylov_1995}. By recognizing
that Krylov methods only need the action of the operator on the vector
instead of the operator itself, the Jacobian can instead be
approximated through various numerical methods including a
difference-based Jacobian-free formulation. Jacobian-Free methods, and
in particular \textit{Jacobian-Free Newton-Krylov} (JFNK) methods
\citep{knoll_jacobian-free_2004}, rely on forming the action of the
Jacobian on a vector as required by the Krylov solver through a
forward difference scheme. In this case, the action of the Jacobian on
some vector $\ve{v}$ is given as:
\begin{equation}
  \ve{J}(\ve{u})\ve{v} = \frac{\ve{F}(\ve{u} + \epsilon \ve{v}) -
    \ve{F}(\ve{u})}{\epsilon}\:,
  \label{eq:jacobian_free_product}
\end{equation}
where $\epsilon$ is a small number typically on the order of machine
precision. Kelley \citep{kelley_iterative_1995} points out a potential
downfall of this formulation in that if the discretization error in
$\ve{F}(\ve{u})$ is on the order of the perturbation parameter
$\epsilon$, then the finite difference error from
Eq~(\ref{eq:jacobian_free_product}) pollutes the solution.

\subsubsection{Automatic Differentiation for Jacobian Generation}
\label{subsubsec:automatic_differentiation}

\section{The FANM Method}
\label{sec:fanm}
In this section we will devise the FANM method. Based on the
background in the previous section, we first need to motivate the
development of this work by pointing out some of the issues with
conventional methods. This includes JFNK coarseness and either
prohibitively large Krylov subspaces or a loss of orthogonality
information in the subspace due to restarts. We can than point out
some of the attractive qualities of the FANM method. Much of this will
rely on advanced concepts including automatic differentiation and
on-the-fly residual and Jacobian generation from nonlinear function
evaluations. In addition, we will want to show that the Newton method
will converge using MCSA as the linear solver. I suspect I will have
some preliminary results for a serial implementation here as I already
have a serial MCSA implemented. Finally, we will want to design the
model problems that we will test FANM with. These should be problems
that have already been worked up in the literature with benchmark-type
solutions. Parallelism in the context of the linear solvers chapter
should also be discussed.

\subsection{Residual and Jacobian Generation}
\label{subsec:fanm_generation}

\subsection{Jacobian Storage vs. Subspace Storage and Restarts}
\label{subsec:fanm_storage}

\subsection{Parallel FANM Implementation}
\label{subsec:parallel_fanm}