### 2012-09-29 doc/prelim/document/chapters/introduction/introduction.tex
adding figure to introduction

\chapter{Introduction}
\label{ch:introduction}
In many fields of engineering and physics, linear and nonlinear
problems are a primary focus of study. Recent focus on multiple
physics systems in the nuclear reactor modeling and simulation
community adds a new level of complexity to common nonlinear systems
as solution strategies change when they are coupled to other problems
\citep{u.s._department_of_energy_casl_2011}. Furthermore, a desire
for predictive simulations to enhance the safety and performance of
nuclear systems creates a need for extremely high fidelity
computations to be performed for these coupled systems as a means to
capture effects not modeled by coarser methods.

In order to achieve this high fidelity, state-of-the-art computing
\index{facilities} must be leveraged in a way that is both efficient and
considerate of hardware-related issues. As scientific computing moves
towards exascale facilities with machines of $O(1,000,000)$ cores
already coming online, new algorithms to solve these complex problems
must be developed to leverage this new hardware
\citep{kogge_using_2011}. Issues such as resiliency to node failure,
limited growth of memory available per node, and scaling to large
numbers of cores will be pertinent to robust algorithms aimed at this
new hardware. Considering these issues, this preliminary report
proposes the development of a massively parallel stochastic method for
linear problems and a novel stochastic method to advance solution
techniques for nonlinear problems.

We discuss in this chapter our motivation for developing these Monte
Carlo solvers by providing background on the current state of solver
techniques for multiphysics problems and the hardware that they
leverage. In addition, multiphysics problems of interest to the
nuclear engineering community are provided to further motivate this
work. We then provide a formal presentation of the multiphysics system
and the problems that arise from it.

\section{Motivation}
\label{sec:motivation}
Predictive modeling and simulation capability requires the combination
of high fidelity models, high performance computing hardware that can
handle the intense compute loads required by these models, and modern
algorithms for solving these problems that leverage this high
performance hardware. For nuclear reactor analysis, this predictive
capability can enable tighter design tolerances for improved thermal
performance and efficiency, higher fuel burnup and therefore reduction
in generated waste, and high confidence in accident scenario
models. The physics that dominate these types of analysis include
neutronics, thermal hydraulics, computational fluid dynamics, and
structural mechanics.

Although solution techniques in each of these individual categories
has advanced over the last few decades and in fact leveraged modern
algorithms and computer architectures, true predictive capability for
engineered systems can only be achieved through a coupled, multiple
physics analysis where the effects of feedback between physics are
modeled. For example, consider the safety analysis of a departure from
nucleate boiling scenario in the subchannel of a nuclear fuel assembly
as shown in Figure~\ref{fig:dnb_example}.
\begin{figure}[htpb!]
  \begin{center}
    \scalebox{1.5}{
      \input{chapters/introduction/dnb_example.pdftex_t} }
  \end{center}
  \caption{\textbf{Multiphysics analysis of departure from nuclear
      boiling.} \textit{A neutronics solution is required to compute
      power generation in the fuel pins, fluid dynamics is required to
      characterize boiling and fluid temperature and density, heat
      transfer is required to compute the fuel and cladding
      temperature, and the nuclear data modified with the temperature
      and density data. Strong coupling among the variables creates a
      nonlinearities.}}
  \label{fig:dnb_example}
\end{figure}
When this event occurs, heat transfer is greatly reduced between the
fuel and the coolant, causing the fuel centerline temperature to
rapidly rise. To characterize this boiling phenomena and how it
affects fuel failure we must consider a neutronics analysis in order
to compute power generation in the fuel pins, fluid dynamics analysis
to characterize coolant boiling, temperature, and density, solid
material heat transfer to characterize fuel and cladding temperature
and heat transfer with the coolant, and nuclear data processing to
characterize how changing material temperatures and densities changes
the cross sections needed for the neutronics calculation. As shown in
Figure~\ref{fig:dnb_example}, many couplings are required among
individual physics components in order to accurately model this
situation with each physics generating and receiving many
responses. Those variables that are very tightly coupled, such as the
temperatures generated by the fluid dynamics and heat transfer
components, will have strong nonlinearities in their behavior and
would therefore benefit from fully consistent nonlinear solution
schemes instead of fixed-point type iterations between
physics\footnote{Fixed-point iterations between physics are commonly
  referred to as Picard iterations.}. Furthermore, the space and time
scales over which these effects occur will also vary greatly.

\section{The Multiphysics System}
\label{sec:multiphysics_system}

\subsection{Multiphysics Methods Advancement}
\label{subsec:methods_advancement}
For some time, the particle transport community has been utilizing
Monte Carlo methods for the solution of transport problems
\citep{lewis_computational_1993}. The partial differential equation
(PDE) community has focused on various deterministic methods for
solutions to linear problems \citep{saad_iterative_2003,
  kelley_iterative_1995}. In between these two areas are a not widely
known small group of stochastic methods for solving sparse linear
systems \citep{forsythe_matrix_1950, hammersley_monte_1964,
  halton_sequential_1962, halton_sequential_1994}. In recent years,
these methods have been further developed for transport problems in
the form of Monte Carlo Synthetic-Acceleration (MCSA)
\citep{evans_residual_2003, evans_monte_2009, evans_monte_2012} but
have yet to be applied to more general sparse linear systems commonly
generated by the computational physics community. Compared to other
methods in this regime, MCSA offers three attractive qualities; (1)
the linear problem operator need not be symmetric or
positive-definite, thereby reducing preconditioning complexity, (2)
the stochastic nature of the solution method provides a natural
solution to the issue of resiliency, and (3) is amenable to
parallelization using modern methods developed by the transport
community \citep{wagner_hybrid_2010}. The developement of MCSA as a
general linear solver and the development of a parallel MCSA method
will be new and unique features of this work, providing a framework
with which other issues such as resiliency may be addressed in the
future.

In addition to linear solver advancements, nonlinear solvers may also
benefit from a general and parallel MCSA scheme. In the nuclear
engineering community, nonlinear problems are often addressed by
either linearizing the problem or building a segregated scheme and
using traditionaly iterative or direct methods to solve the resulting
system \citep{pletcher_computational_1997}. In the mathematics
community, various Newton methods have been popular
\citep{kelley_iterative_1995}. Recently, Jacobian-Free Newton-Krylov
(JFNK) schemes \citep{knoll_jacobian-free_2004} have been utilized in
multiple physics architectures and advanced single physics codes
\citep{gaston_parallel_2009}. The benefits of JFNK schemes are that
the Jacobian is never formed, simplifying the implementation, and a
Krylov solver is leveraged (typically GMRES or Conjugate Gradient),
providing excellent convergence properties for well-conditioned and
well-scaled systems. However, there are two potential drawbacks to
these methods for high fidelity predictive simulations: (1) the
Jacobian is approximated by a first-order differencing method on the
order of machine precision such that this error can grow beyond that
of those in a fine-grained system \citep{kelley_iterative_1995} and
(2) for systems that are not symmetric positive-definite (which will
be the case for most multiphysics systems and certainly for most
preconditioned systems) the Krylov subspace generated by the GMRES
solver may become prohibitively large
\citep{knoll_newton-krylov_1995}. To address these issues, this thesis
proposes a new and novel method for nonlinear systems based on the
MCSA method.

The Forward-Automated Newton-MCSA (FANM) method is proposed as new
nonlinear solution method. The key features of FANM are: full Jacobian
generation using modern Forward Automated Differencing (FAD) methods
\citep{bartlett_automatic_2006}, and MCSA as the inner linear
solver. This method has several attractive properties. First, the
first-order approximation to the Jacobian used in JFNK type methods is
eliminated by generating the Jacobian explicitly with the model
equations through FAD. Second, the Jacobian need not be explicitly
formed by the user but is instead automated through FAD; this
eleminates the complexity of hand-coding derivatives and has also been
demonstrated to be more efficient computationally than evaluating
differenced derivatives. Third, unlike GMRES, MCSA does not build a
subspace during iterations. Although the Jacobian must be explicitly
formed to use MCSA, for problems that take more than a few GMRES
iterations to converge the size of the Krylov subspace will grow
beyond that of the Jacobian. Finally, using MCSA for the linear solve
provides its benefits for preconditioning, potential resiliency, and
parallelism.

\section{Research Outline}
\label{sec:research_outline}
This preliminary report oultines the conventional methods used in
practice for solving linear and nonlinear problems to provide a
mathematical basis upon which to build new algorithms that aim to
solve some of the aforementioned issues. These new algorithms will be
outlined in full with links made to past work and their potential for
offering improvements to the computational physics community. Progress
to date will be reported including the development of a new physics
framework that will serve as a test bed for these new methods that
leverages a preliminary implementation of the stochastic method. A
research plan will then be provided that continues the development and
implementation of these new methods and verifies them with benchmark
solutions, culminating in the analysis of a multiphysics simulation of
a pressurized water reactor assembly.