### 2012-12-10 doc/sna2013/abstract/sna_abstract.tex
adding makefile to mandc 2013

\documentclass[letterpaper,12pt]{article}
\usepackage[top=1.0in,bottom=1.0in,left=1.25in,right=1.25in]{geometry}
\usepackage{verbatim}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{tabls}
\usepackage{afterpage}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[mathcal]{euscript}
\usepackage[usenames]{color}
\usepackage[
naturalnames = true, 
colorlinks = true, 
linkcolor = black,
anchorcolor = black,
citecolor = black,
menucolor = black,
urlcolor = blue
]{hyperref}

%%---------------------------------------------------------------------------%%
\author{Stuart R. Slattery, Thomas M. Evans, Paul P.H. Wilson
\\ \href{mailto:sslattery@wisc.edu}{\texttt{sslattery@wisc.edu}}
}

\date{\today} 
\title{A Multiple-Set Overlapping-Domain Decomposed Monte Carlo
  Synthetic Acceleration Method for Linear Systems} 
\begin{document}
\maketitle

\abstract
{
We present a novel multiple-set overlapping-domain decomposed strategy
for parallelizing the Monte Carlo Synthetic Acceleration (MCSA)
method. MCSA methods use the Neumann-Ulam class of Monte Carlo solvers
for linear systems to accelerate the fixed point method upon which
the Neumann-Ulam solvers are based. To effectively parallelize MCSA
methods requires the parallelization of the Neumann-Ulam solvers. To
do this in a domain decomposed environment, we borrow strategies
traditionally implemented in reactor physics to parallelize the
problem. The multiple-set overlapping-domain decomposition algorithm
is presented along with preliminary scaling results.
}

%%---------------------------------------------------------------------------%%
\section{Introduction}
As leadership class machines move towards the exascale, new algorithms
must be developed that leverage their strengths and adapt to their
shortcomings. Basic research is required now to advance methods in
time for these new machines to become operational. We can identify key
issues for which a set of robust, massively parallel Monte Carlo
solvers could provide a solution. As machines begin to operate at
hundreds of petaflops peak performance and beyond, trends toward
reduced energy consumption will require incredibly high levels of
concurrency to achieve the desired computation rates. Furthermore,
this drop in power consumption will mean increased pressure on memory
as memory per node is expected to stagnate while cores per node is
expected to increase. As the number of cores increases, their clock
speed is expected to stagnate or even decrease to further reduce power
consumption and manufacturing costs.

The end result of these hardware changes is that the larger numbers of
low-powered processors will be prone to both soft failures such as bit
errors in floating point operations and hard failures where the data
owned by that processor cannot be recovered. Because these failures
are predicted to be common, resilient solver technologies are required
to overcome these events. With linear solvers based on Monte Carlo
techniques, such issues are alleviated by statistical arguments. In
the case of soft failures, isolated floating point errors in Monte
Carlo simulation are absorbed within tally statistics while completely
losing hardware during a hard failure is manifested as a high variance
event where some portion of the Monte Carlo histories are lost.

%%---------------------------------------------------------------------------%%
\section{Monte Carlo Synthetic Acceleration}

An alternative approach to approximate matrix inversion is to employ
Monte Carlo methods that sample a distribution with an expectation
value equivalent to that of the inverted operator. Such methods have
been in existence for decades with the earliest reference noted here
an enjoyable manuscript published in 1950 by Forsythe and Leibler
\cite{forsythe_matrix_1950}. In their outline, Forsythe and Liebler
in fact credit the creation of this technique to J. Von Neumann and
S.M. Ulam some years earlier than its publication. In 1952 Wasow
provided a more formal explanation of Von Neumann and Ulam's method
\cite{wasow_note_1952} and Hammersley and Handscomb's 1964 monograph
\cite{hammersley_monte_1964} and Spanier and Gelbard's 1969 book
\cite{spanier_monte_1969} present additional detail on this topic
using a collection of references from the 1950's and early 1960's.

Using the ideas of Halton, Evans and Mosher recently developed a Monte
Carlo solution method that was not prohibited severely by the quality
of the initial guess for the system \cite{evans_monte_2009} and later
applied it more rigorously as a solution mechanism for the radiation
diffusion equation \cite{evans_monte_2012}. With their new methods,
they achieved identical numerical results as and marginally better
performance than conventional Krylov solvers. Their approach was
instead to use residual Monte Carlo as a synthetic acceleration for a
stationary method. The \textit{Fixed-Point Monte Carlo
  Synthetic-Acceleration} (MCSA) method is defined as:
\begin{subequations}
  \begin{gather}
    \ve{x}^{k+1/2} = (\ve{I} - \ve{A})\ve{x}^k + \ve{b}\:,\\
    \ve{r}^{k+1/2} = \ve{b} - \ve{A}\ve{x}^{k+1/2}\:,\\
    \ve{A}\delta\ve{x}^{k+1/2} = \ve{r}^{k+1/2}\:,\\
    \ve{x}^{k+1} = \ve{x}^{k+1/2} + \delta \ve{x}^{k+1/2}\:,
  \end{gather}
  \label{eq:mcsa}
\end{subequations}
where the adjoint Monte Carlo method is used to generate the solution
correction from the residual. Using Monte Carlo in this way achieves
the same effect as Halton's method, decoupling its convergence rate
from the overall convergence rate of the method. Here, the approximate
Monte Carlo solution is not driven to a particular convergence as it
merely supplies a correction for the initial guess generated by
Richardson's iteration. Rather, only a set number of histories are
required using the adjoint method to generate the correction.

%%---------------------------------------------------------------------------%%
\section{Multiple-Set Overlapping-Domain Decomposition}

In 2010, Wagner and colleagues developed the \textit{multiple-set
  overlapping-domain} (MSOD) decomposition for parallel Monte Carlo
applications for full-core light water reactor analysis
\cite{wagner_hybrid_2010}. In their work, an extension of Brunner's,
their scheme employed the similar parallel algorithms for particle
transport but a certain amount of overlap between adjacent domains was
used to decrease the number of particles leaving the local domain. In
addition, Wagner utilized a level of replication of the domain such
that the domain was only decomposed on $O(100)$ processors and if
replicated $O(1,000)$ times achieves simulation on $O(100,000)$
processors, thus providing spatial and particle parallelism. Each
collection of processors that constitutes a representation of the
entire domain is referred to as a set, and within a set overlap occurs
among its sub-domains. The original motivation was to decompose the
domain in a way that it remained in a physical cabinet in a large
distributed machine, thus reducing latency costs during
communication. A multiple set scheme is also motivated by the fact
that communication during particle transport only occurs within a set,
limiting communications during the transport procedure to a group of
$O(100)$ processors, a number that was shown to have excellent
parallel efficiencies in Brunner's work and therefore will scale well
in this algorithm. The overlapping domains within each set also
demonstrated reduced communication costs. On each processor, the
source is sampled in the local domain that would exist if no overlap
was used while tallies can be made over the entire overlapping domain.

%%---------------------------------------------------------------------------%%
\section{Results}

%%---------------------------------------------------------------------------%%
\section{Conclusion}

%%---------------------------------------------------------------------------%%
\pagebreak
\bibliographystyle{ieeetr}
\bibliography{paper}
\end{document}