### 2013-02-27 doc/spn/spn.tex
finished Pn derivation, boundary conditions are next

%        File: spn.tex
\documentclass[letterpaper]{article}
\usepackage[top=1.0in,bottom=1.0in,left=1.0in,right=1.0in]{geometry}
\usepackage{verbatim}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage[usenames]{color}
\usepackage[
naturalnames = true, 
colorlinks = true, 
linkcolor = black,
anchorcolor = black,
citecolor = black,
menucolor = black,
urlcolor = blue
]{hyperref}
\usepackage{listings}
\usepackage{textcomp}
\definecolor{listinggray}{gray}{0.9}
\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}
\lstset{
  backgroundcolor=\color{lbcolor},
  tabsize=4,
  rulecolor=,
  language=c++,
  basicstyle=\scriptsize,
  upquote=true,
  aboveskip={1.5\baselineskip},
  columns=fixed,
  showstringspaces=false,
  extendedchars=true,
  breaklines=true,
  prebreak =
  \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
  frame=single,
  showtabs=false,
  showspaces=false,
  showstringspaces=false,
  identifierstyle=\ttfamily,
  keywordstyle=\color[rgb]{0,0,1},
  commentstyle=\color[rgb]{0.133,0.545,0.133},
  stringstyle=\color[rgb]{0.627,0.126,0.941},
}

\author{Stuart R. Slattery
  \\ \href{mailto:sslattery@wisc.edu}{\texttt{sslattery@wisc.edu}}
}

\date{\today}
\title{Monte Carlo Solutions for the $SP_N$ Equations}
\begin{document}
\maketitle

%%---------------------------------------------------------------------------%%
\section{Introduction}
\label{sec:introduction}

%%---------------------------------------------------------------------------%%
\section{Derivation of the $SP_N$ Equations}
\label{sec:derivation}
In this section, we derive the $SP_N$ equations. We begin by stating
the general time-independent neutron transport equation followed by a
derivation of the $P_N$ equations in planar geometry for multiple
energy groups. From these equations, we then apply a set of
approximations to yield the multi-dimensional, multi-group $SP_N$
equations for fixed source problemsd

\subsection{The Neutron Transport Equation}
\label{subsec:transport_eq}
As a starting point we define the time-independent neutron transport
equation \cite{lewis_computational_1993}:
\begin{equation}
  \hat{\Omega} \cdot \vec{\nabla} \psi(\vec{r},\hat{\Omega},E) +
  \sigma(\vec{r},E) \psi(\vec{r},\hat{\Omega},E) = \int dE' \int
  d\Omega' \sigma_s(\vec{r},E' \rightarrow E,\hat{\Omega}' \rightarrow
  \hat{\Omega}) \psi(\vec{r},\hat{\Omega}',E') +
  q(\vec{r},\hat{\Omega},E)\:,
  \label{eq:general_transport}
\end{equation}
with the variables defined as:
\begin{itemize}
\item $\vec{r}$ - neutron spatial position
\item $\hat{\Omega}$ - neutron streaming direction with radial
  component $\mu$ and azimuthal component $\omega$
\item $\hat{\Omega}' \cdot \hat{\Omega} = \mu_0$ is the angle of
  scattering
\item $E$ - neutron energy
\item $\psi(\vec{r},\hat{\Omega},E)$ - angular flux
\item $\sigma(\vec{r},E)$ - total interaction cross section
\item $\sigma_s(\vec{r},E' \rightarrow E,\hat{\Omega}'$ - probability
  of scattering from direction $\hat{\Omega}'$ into an angular domain
  $d\hat{\Omega}'$ about the direction $\hat{\Omega}$ and from energy
  $E'$ to an energy domain $dE'$ about energy $E$
\item $q(\vec{r},\hat{\Omega},E)$ - external source of neutrons.
\end{itemize}
For this work, it is sufficient to formulate
Eq~(\ref{eq:general_transport}) in 1-dimensonal Cartesian geometry:
\begin{equation}
  \mu \frac{\partial}{\partial x} \psi(x,\mu,E) + \sigma(x,E)
  \psi(x,\mu,E) = \int dE' \int d\Omega' \sigma_s(x,E' \rightarrow
  E,\hat{\Omega}' \rightarrow \hat{\Omega})
  \psi(\vec{r},\hat{\Omega}',E') + \frac{q(x,E)}{4 \pi}\:,
  \label{eq:cart_1d_transport}
\end{equation}
where the angular component of the solution is no longer dependent on
the azimuthal direction of travel and an isotropic source of neutrons
is assumed.

\subsection{Derivation of the $P_N$ Equations}
\label{subsec:pn_equations}
Next, we derive the $P_N$ equations, a simplified form of the general
transport equation where Legendre polynomials are used to expand the
angular flux and scattering cross section variables as a means of
capturing the angular structure of the solution. Before deriving this
form of the transport equation, we briefly discuss a few properties of
Legendre polynomials that we will find useful in the derivation.

\subsubsection{Legendre Polynomials}
\label{subsubsec:legendre_polys}
The Legendre polynomials are an orthogonal set of functions that
are solutions to Legendre's differential equation. They have the
following form \cite{lewis_computational_1993}:
\begin{equation}
  P_l(\mu) = \frac{1}{2^l l!}\frac{d^l}{d \mu^l}(\mu^2-1)^l\:.
  \label{eq:general_legendre_poly}
\end{equation}
These functions have several nice properties including orthogonality:
\begin{equation}
  \int_{-1}^{1} P_l(\mu) P_{l'}(\mu) d\mu = \frac{1}{2l+1}\delta_{l l'}\:,
  \label{eq:legendre_orthog}
\end{equation}
a recurrence relation:
\begin{equation}
  \mu P_l(\mu) = \frac{1}{2l+1}[(l+1)P_{l+1}(\mu) + l P_{l-1}(\mu)]\:,
  \label{eq:legendre_recurrence}
\end{equation}
and an addition theorem:
\begin{equation}
  P_l(\hat{\Omega} \cdot \hat{\Omega}') = \frac{1}{2n+1}\sum_{m=-l}^l
  Y_{lm}(\hat{\Omega})Y^*_{lm}(\hat{\Omega}')\:.
  \label{eq:legendre_addition}
\end{equation}
where the functions $Y_{lm}(\hat{\Omega})$ are the spherical
harmonics. We can form the addition theorem in this way because the
spherical harmonics are in fact just harmonic multiples of the
Legendre polynomials:
\begin{equation}
  Y_{lm}(\hat{\Omega}) =
  \sqrt{\frac{(2l+1)(l-m)!}{(l+m)!}}P^m_l(\mu)e^{i \omega m}\:,
  \label{eq:spherical_harmonic}
\end{equation}
where $\omega$ is the azimuthal component of the streaming
direction. We can reduce Eq~(\ref{eq:legendre_addition}) for the
planar geometry we are studying by ignoring the azimuthal components
of the addition theorem. As shown in Eq~(\ref{eq:spherical_harmonic}),
the azimuthal dependence is given by the harmonic component, $e^{i
  \omega m}$, and therefore we choose to ignore all terms in
Eq.~(\ref{eq:legendre_addition}) where $m \neq 0$. This gives:
\begin{equation}
  P_l(\hat{\Omega} \cdot \hat{\Omega}') = \frac{1}{2n+1}
  Y_{l0}(\hat{\Omega})Y^*_{l0}(\hat{\Omega}')\:.
  \label{eq:legendre_addition_2}
\end{equation}
Per Eq~(\ref{eq:spherical_harmonic}) we have:
\begin{equation}
  Y_{l0}(\hat{\Omega}) = \sqrt{2l+1}P_l^0(\mu)\:,
  \label{eq:harmonic_0}
\end{equation}
where $P_l^0(\mu) = P_l(\mu)$ is the $0^{th}$ associated Legendre
function. Finally, we can reduce the addition theorem from
Eq~(\ref{eq:legendre_addition_2}) with Eq~(\ref{eq:harmonic_0}) to a
simple product for planar geometry:
\begin{equation}
  P_l(\hat{\Omega} \cdot \hat{\Omega}') = P_l(\mu)P_l(\mu')\:.
  \label{eq:legendre_addition_3}
\end{equation}

\subsubsection{Planar $P_N$ Equations}
\label{subsubsec:planar_pn}
With the Legendre polynomial properties defined above, we can proceed
by deriving the $P_N$ equations for planar geometry. For these
equations, we will start by assuming a monoenergtic field of neutrons
such that we are solving the following reduced form of the transport
equation: 
\begin{equation}
  \mu \frac{\partial}{\partial x} \psi(x,\mu) + \sigma(x) \psi(x,\mu)
  = \int d\Omega' \sigma_s(x,\hat{\Omega}' \rightarrow \hat{\Omega})
  \psi(\vec{r},\hat{\Omega}') + \frac{q(x)}{4 \pi}\:.
  \label{eq:mono_transport}
\end{equation}
The $P_N$ equations introduce the approximation that the angular
dependence of the scattering cross section, $sigma_s$, and the angular
flux $\psi$, can be discretized by expanding them in Legendre
polynomials as follows:
\begin{equation}
  \psi(x,\mu) = \sum_{n=0}^\infty (2n+1)P_n(\mu)\phi_n(x)\:,
  \label{eq:flux_expansion}
\end{equation}
\begin{equation}
  \sigma_{sm}(x) = \sum_{m=0}^\infty (2m+1)P_m(\mu)\sigma_s(x)\:,
  \label{eq:scattering_expansion}
\end{equation}
where we have supressed the $2\pi$ generated by integrating away the
azimuthal angular component and $\phi_n(x)$ in
Eq~(\ref{eq:flux_expansion}) is referred to as the $n^{th}$ Legendre
moment of the neutron flux and is given by:
\begin{equation}
  \phi_n(x) = \int_{-1}^1 P_n(\mu)\psi(x,\mu)d\mu\:.
  \label{eq:legendre_moments}
\end{equation}
We first insert the expansions given by Eq~(\ref{eq:flux_expansion})
and (\ref{eq:scattering_expansion}) into the planar transport equation
given by Eq~(\ref{eq:mono_transport}):
\begin{equation}
  \frac{\partial}{\partial x}\Big[\sum_{n=0}^\infty (2n+1) \phi_n \mu
    P_n(\mu) \Big] + \sigma \sum_{n=0}^\infty (2n+1) \phi_n P_n(\mu) =
  \int_{-1}^1 \sum_{m=0}^\infty (2m+1) \sigma_{sm} P_m(\mu_0)
  \sum_{n=0}^\infty (2n+1) \phi_n P_n(\mu') d \mu' + q\:,
  \label{eq:pn_deriv_1}
\end{equation}
where the dependence on the spatial variable, $x$, has been
supressed. To arrive at the $P_N$ equations we multiply
Eq~(\ref{eq:pn_deriv_1}) by $P_m(\mu)$ and integrate over the angular
domain $\int_{-1}^1 d \mu$. We will look at each term in
Eq~(\ref{eq:pn_deriv_1}) individually.

\paragraph{Streaming Term}
We first apply the multiplication and integration as prescribed above:
\begin{equation}
  \frac{\partial}{\partial x}\Bigg[\sum_{n=0}^\infty (2n+1) \mu \phi_n
    P_n(\mu) \Bigg] \rightarrow \int_{-1}^1 \frac{\partial}{\partial
    x}\Bigg[\sum_{n=0}^\infty (2n+1) \phi_n \mu P_n(\mu) P_m(\mu) \Bigg]
  d \mu\:.
  \label{eq:pn_deriv_2}
\end{equation}
The $\mu P_n(\mu)$ term can be eliminated via the recurrence relation
given by Eq~(\ref{eq:legendre_recurrence}):
\begin{equation}
\int_{-1}^1 \frac{\partial}{\partial x}\Bigg[\sum_{n=0}^\infty (2n+1)
  \frac{\phi_n}{2n+1}[(n+1)P_{n+1}(\mu) + n P_{n-1}(\mu)] P_m(\mu)
  \Bigg] d\mu\:,
\label{eq:pn_deriv_3}
\end{equation}
which expands to:
\begin{equation}
  \sum_{n=0}^\infty \frac{\partial}{\partial
    x}\phi_n\Bigg[(n+1)\int_{-1}^1 P_{n+1}(\mu)P_m(\mu) d\mu + n
    \int_{-1}^1 P_{n-1}(\mu) P_m(\mu) d\mu \Bigg] \:.
  \label{eq:pn_deriv_4}
\end{equation}
This reveals the orthogonality relation given by
Eq~(\ref{eq:legendre_orthog}) that when inserted into
Eq~(\ref{eq:pn_deriv_3}):
\begin{equation}
  \sum_{n=0}^\infty \frac{\partial}{\partial
    x}\phi_n\Bigg[(n+1)\frac{1}{2n+1}\delta_{n,n+1} +
    n\frac{1}{2n+1}\delta_{n,n-1} \Bigg] \:.
  \label{eq:pn_deriv_5}
\end{equation}
We can then distribute the Legendre moment to arrive at the final form
of the streaming term:
\begin{equation}
  \sum_{n=0}^\infty \frac{\partial}{\partial x} \frac{1}{n+1} \Big[
    (n+1) \phi_{n+1} + n \phi_{n-1} \Big] \:.
  \label{eq:pn_deriv_6}
\end{equation}

\paragraph{Removal Term}
To reduce the removal term, the orthogonality relation is again
applied after the integration:
\begin{equation}
  \sigma \sum_{n=0}^\infty (2n+1) \phi_n P_n(\mu) \rightarrow
  \int_{-1}^1 \sigma \sum_{n=0}^\infty (2n+1) \phi_n P_n(\mu) P_m(\mu)
  d\mu\:,
  \label{eq:pn_deriv_7}
\end{equation}
\begin{equation}
  \sigma \sum_{n=0}^\infty (2n+1) \phi_n \frac{1}{2n+1}\:,
  \label{eq:pn_deriv_8}
\end{equation}
giving for the final removal term:
\begin{equation}
  \sum_{n=0}^\infty \sigma \phi_n \:.
  \label{eq:pn_deriv_9}
\end{equation}

\paragraph{Scattering Term}
For the scattering term:
\begin{equation}
  \int_{-1}^1 \sum_{m=0}^\infty (2m+1) \sigma_{sm} P_m(\mu_0)
  \sum_{n=0}^\infty (2n+1) \phi_n P_n(\mu') d\mu' \rightarrow
  \int_{-1}^1 \int_{-1}^1 \sum_{m=0}^\infty (2m+1) \sigma_{sm} P_m(\mu)
  P_m(\mu_0) \sum_{n=0}^\infty (2n+1) \phi_n P_n(\mu') d\mu' d\mu\:.
  \label{eq:pn_deriv_10}
\end{equation}
The addition thereom from Eq~(\ref{eq:legendre_addition_3}) is applied
to give:
\begin{equation}
  \int_{-1}^1 \int_{-1}^1 \sum_{m=0}^\infty (2m+1) \sigma_{sm} P_m(\mu)
  P_m(\mu)P_m(\mu') \sum_{n=0}^\infty (2n+1) \phi_n P_n(\mu') d\mu' d\mu\:,
  \label{eq:pn_deriv_11}
\end{equation}
which can be rearranged as:
\begin{equation}
  \sum_{m=0}^\infty \sum_{n=0}^\infty (2m+1) \sigma_{sm} (2n+1) \phi_n
  \int_{-1}^1 P_m(\mu') P_n(\mu') d\mu' \int_{-1}^1 P_m(\mu) P_m(\mu)
  d\mu\:.
  \label{eq:pn_deriv_12}
\end{equation}
Again we can apply orthogonality to eliminate the Legendre polynomials:
\begin{equation}
  \sum_{m=0}^\infty \sum_{n=0}^\infty (2m+1) \sigma_{sm} (2n+1) \phi_n
  \frac{1}{2n+1}\delta_{nm}\frac{1}{2m+1}\delta_{mm}\:,
  \label{eq:pn_deriv_13}
\end{equation}
which is reduced to:
\begin{equation}
  \sum_{n=0}^\infty \sigma_{sn} \phi_n\:,
  \label{eq:pn_deriv_14}
\end{equation}
with the dependence on the index $m$ eliminated.

\paragraph{Source Term}
The laste term we are concerned with in Eq~(\ref{eq:pn_deriv_1}) is
the source term:
\begin{equation}
  q \rightarrow \int_{-1}^1 q P_n(\mu) d\mu \:.
  \label{eq:pn_deriv_15}
\end{equation}
We can leverage orthogonality by multiplying by $P_0(\mu) = 1$:
\begin{equation}
  \int_{-1}^1 q P_0 P_n(\mu) d\mu = \frac{q}{2*0+1}\delta_{n0}\:,
  \label{eq:pn_deriv_16}
\end{equation}
giving a final source term of:
\begin{equation}
  q\delta_{n0}\:.
  \label{eq:pn_deriv_17}
\end{equation}

Now that we have expanded all angular dependent terms in
Eq~(\ref{eq:pn_deriv_1}) and reduced them appropriately, we can
combine them to generate the $P_N$ equations:
\begin{equation}
    \sum_{n=0}^\infty \frac{\partial}{\partial x} \frac{1}{n+1} \Big[
      (n+1) \phi_{n+1} + n \phi_{n-1} \Big] + \sum_{n=0}^\infty \sigma
    \phi_n = \sum_{n=0}^\infty \sigma_{sn} \phi_n + q\delta_{n0}\:.
  \label{eq:pn_deriv_18}
\end{equation}
More formally, the $P_N$ equations are written as:
\begin{equation}
   \frac{1}{n+1} \frac{\partial}{\partial x}\Big[ (n+1) \phi_{n+1} + n
     \phi_{n-1} \Big] + \Sigma_n \phi_n = q\delta_{n0} \:,
  \label{eq:final_pn_equations}
\end{equation}
where $\Sigma_n = \sigma-\sigma_{sn}$ and the summations are truncated
at some level of approximation such that $n = 0,1,\dotsc,N$. This
yields a set of $N+1$ equations for $N+2$ flux moments. We therefore
require an additional equation to close the system. In accordance with
the series truncation as an approximation we choose:
\begin{equation}
  \phi_{N+1} = 0\:.
  \label{eq:pn_closure}
\end{equation}
As an example, we will construct the P5 equations from
Eq~(\ref{eq:final_pn_equations}) and the closure given by
Eq~(\ref{eq:pn_closure}):
\begin{subequations}
  \begin{gather}
   \frac{\partial}{\partial x}\phi_{1} + \Sigma_0 \phi_0 = q\:,\\ 
   \frac{1}{2} \frac{\partial}{\partial x}\Big[ 2
     \phi_{2} + \phi_{0} \Big] + \Sigma_1 \phi_1 = 0\:,\\
   \frac{1}{3} \frac{\partial}{\partial x}\Big[ 3 \phi_{3} + 2
     \phi_{1} \Big] + \Sigma_2 \phi_2 = 0 \:,\\
   \frac{1}{4} \frac{\partial}{\partial x}\Big[ 4 \phi_{4} + 3
     \phi_{2} \Big] + \Sigma_3 \phi_3 = 0 \:,\\
   \frac{1}{5} \frac{\partial}{\partial x}\Big[ 5 \phi_{5} + 4
     \phi_{3} \Big] + \Sigma_4 \phi_4 = 0 \:,\\
   \frac{1}{n+1} \frac{\partial}{\partial x} 5 \phi_{4} + \Sigma_5
   \phi_5 = 0 \:.
  \end{gather}
  \label{eq:p5_equations}
\end{subequations}
This gives us a set of 6 coupled equations for the six Legendre
moments requested defined over the entire spatial domain for a single
energy group.

\subsubsection{Boundary Conditions for the $P_N$ Equations}
\label{subsbusec:bcs_pn}

\subsection{Formulation of the  $SP_N$ Equations}
\label{subsec:spn_equations}

%%---------------------------------------------------------------------------%%
\section{Spectral Analysis of the $SP_N$ Equations}
\label{sec:spectral_analysis}

%%---------------------------------------------------------------------------%%
\section{Monte Carlo Solutions for the $SP_N$ Equations}
\label{sec:monte_carlo}

%%---------------------------------------------------------------------------%%
\section{Conclusion}
\label{sec:conclusion}

%%---------------------------------------------------------------------------%%
\pagebreak
\bibliographystyle{ieeetr} \bibliography{references}
\end{document}