### 2013-03-21 doc/dissertation/document/frontmatter/abstract.tex
adding dissertation document. might as well start now

\svnidlong{$LastChangedBy$}{$LastChangedRevision$}{$LastChangedDate$}{$HeadURL: http://freevariable.com/dissertation/branches/diss-template/frontmatter/abstract.tex $}
\vcinfo{}