### 2013-03-21 doc/dissertation/document/chapters/navier_stokes_solutions/navier_stokes_solutions.tex
spn derivation dump

\chapter{Monte Carlo Solution Methods for the Navier-Stokes Equations}
\label{ch:mc_ns_solutions}

Put Navier Stokes work here.