### 2013-03-26 doc/dissertation/document/chapters/spn_solutions/spn_solutions.tex
adding richardson iteration matrix spectral radius function

\chapter{Monte Carlo Solution Methods for the Simplified $P_N$ Equations}
\label{ch:mc_spn_solutions}
In this chapter we apply a preconditioned MCSA method as a solution
scheme for the $SP_N$ equations.

%%---------------------------------------------------------------------------%%
\section{Fuel Assembly Scaling Results for the $SP_N$ Equations}
\label{sec:spn_assembly_scaling}

%%---------------------------------------------------------------------------%%
\section{Full Core Scaling Results for the $SP_N$ Equations}
\label{sec:spn_assembly_scaling}