### 2013-05-03 doc/dissertation/document/backmatter/spn_spatial_discretization.tex
adding appending for spn finite volume discretization

\chapter{Finite Volume Discretization for the $SP_N$ Equations}
\label{chap:spn_spatial_discretization}
Finite volume discretization for the $SP_N$ equations.