### 2013-05-16 doc/jcomp2012/response.txt
Final edits.

Dear Editor,

     Thank you for the thorough reviews to our paper.  We hope that we have
     successfully responded to each reviewer.  Our responses follow below:

     Response to reviewer 2.

     We have added an acknowledgement indicating that the early part of the
     work was performed at LANL.  We have added Section 2.4 that does detailed
     comparisons between sequential Monte Carlo and MCSA.  Furthermore, we
     have added text in the discussion that highlights the differences between
     the methods.

     This reviewer made a number of key editorial observations.  We have
     adapted the text for each of them.  Some key points:

     - we added a reference to synthetic acceleration
     - we defined the terms in Eq. 6 and added some text for clarity
     - we didn't add "adjoint" to the abstract because we feel it would be
       confusing to the traditional Monte Carlo transport community.  We feel
       the discussion in the text clarifies the differences sufficiently. We
       also clarified the discussion involving the Kronecker delta in Eq. 13.
     - we chose Np and Wc for the multimaterial problem using numerical
       experiments; we added text indicating this.

     Response to reviewer 3.

     This reviewer correctly observed that the solvers that we compared MCSA
     to in our original manuscript and the way the information was presented
     was deficient.  We have expanded the results section to include GMRES and
     CG results with a family of state-of-the-art preconditioners.  We also
     updated the tables to include average and total iterations along with
     other useful quantities that makes the comparisons more meaningful.
     Additional text has been added to the results section to clarify the
     comparisons.

     We added a sentence clarifying the use of +|- in Eq 51.

     We added a paragraph providing more motivation to the introduction.

     We presented Figure 11 showing the spectral radius as a function of
     problem time in order to address the issue of variable time-step size
     versus solver performance.  Because all solvers use the same time-steps,
     no additional parameterization studies should be required.  We do
     appreciate that different solvers can behave differently based on
     time-step size, we feel that for this model equation, we are capturing
     the comparison between each method fairly.