### 2013-05-16 doc/jcomp2012/response.txt
Final, final edits.

Dear Editor,

     Thank you for the thorough reviews to our paper.  Both reviewers did a
     very thorough job, and in general, we have no objections to any of their
     specific comments.  We believe we have updated the paper to reflect that
     main issues presented by the reviewers.  Other than editorial additions,
     the main items added were:

     - a new section directly comparing sequential MC with MCSA
     - we implemented GMRES and Multigrid preconditioned CG along with ILU and
       ILUT preconditioning and compared these to MCSA for the model problem

     We feel that these improvements address each reviewers primary critique.
     We have endeavered to address each of the specific comments by the
     reviewers, these are addressed individually below:

     Response to reviewer 2.

     We have added an acknowledgement indicating that the early part of the
     work was performed at LANL.  We have added Section 2.4 that does detailed
     comparisons between sequential Monte Carlo and MCSA.  Furthermore, we
     have added text in the discussion that highlights the differences between
     the methods.

     This reviewer made a number of key editorial observations.  We have
     adapted the text for each of them.  Some key points:

     - we added a reference to synthetic acceleration
     - we defined the terms in Eq. 6 and added some text for clarity
     - we didn't add "adjoint" to the abstract because we feel it would be
       confusing to the traditional Monte Carlo transport community.  We feel
       the discussion in the text clarifies the differences sufficiently. We
       also clarified the discussion involving the Kronecker delta in Eq. 13.
     - we chose Np and Wc for the multimaterial problem using numerical
       experiments; we added text indicating this.

     Response to reviewer 3.

     This reviewer correctly observed that the solvers that we compared MCSA
     to in our original manuscript and the way the information was presented
     was deficient.  We have expanded the results section to include GMRES and
     CG results with a family of state-of-the-art preconditioners, including
     algebraic multigrid.  We also updated the tables to include average and
     total iterations along with other useful quantities that makes the
     comparisons more meaningful.  Additional text has been added to the
     results section to clarify the comparisons.

     We added a sentence clarifying the use of +|- in Eq 51.

     We added a paragraph providing more motivation to the introduction.

     We presented Figure 11 showing the spectral radius as a function of
     problem time in order to address the issue of variable time-step size
     versus solver performance.  Because all solvers use the same time-steps,
     no additional parameterization studies should be required.  We do
     appreciate that different solvers can behave differently based on
     time-step size, we feel that for this model equation, we are capturing
     the comparison between each method fairly.