### 2013-07-11 doc/dissertation/document/chapters/conclusion/conclusion.tex
updating preliminaries of nonlinear chapter

\chapter{Conclusions and Analysis\ }
\label{ch:conclusion}
Put conclusions here.

%%---------------------------------------------------------------------------%%
\section{Monte Carlo Synthetic Acceleration Solutions for the $SP_N$ Equations\ }
\label{sec:spn_conclusion}

%%---------------------------------------------------------------------------%%
\section{Monte Carlo Synthetic Acceleration Solutions for Navier-Stokes Equations\ }
\label{sec:nonlinear_conclusions}

%%---------------------------------------------------------------------------%%
\section{Parallel Monte Carlo Synthetic Acceleration\ }
\label{sec:parallel_mc_conclusions}

%%---------------------------------------------------------------------------%%
\section{Future Work\ }
\label{sec:future_work}

Breaking away from $\rho(H) < 1$

Monte Carlo methods of the 2nd Degree?

Stochastic projection methods?

Stochastic variants of conventional preconditioners?

Variance reduction for preconditioning?

Controling MCSA histories as a function of nonlinear iteration/forcing
term?

Implementation optimization - can we approach Krylov performance?

Reduction in memory footprint

Multiple sets parallel overhead reduction

%%---------------------------------------------------------------------------%%
\section{Closing Remarks\ }
\label{sec:closing}