### 2013-07-24 doc/dissertation/document/frontmatter/acks.tex

This work was performed under appointment to the Nuclear Regulatory
Commission Fellowship program at the University of Wisconsin - Madison
Department of Engineering Physics.

Thank you to the Oak Ridge Leadership Computing Facility (OLCF) for
providing computational resources and support.