### 2013-08-20 doc/dissertation/document/frontmatter/acks.tex
adding proper olcf acknowledgement

This work was performed under appointment to the Nuclear Regulatory
Commission Fellowship program at the University of Wisconsin - Madison
Department of Engineering Physics.

This research used resources of the Oak Ridge Leadership Computing
Facility at the Oak Ridge National Laboratory, which is supported by
the Office of Science of the U.S. Department of Energy under Contract
No. DE-AC05-00OR22725.