### 2013-08-22 doc/dissertation/document/chapters/nonlinear_problem/nonlinear_problem.tex
removing redundant fanm verification data.

\chapter{Monte Carlo Synthetic Acceleration\\ Methods for the
  Navier-Stokes\\ Equations\ }
\label{ch:nonlinear_problem}

In this chapter the fundamentals of Newton's method will be presented
along with a brief discussion of inexact Newton methods. Newton-Krylov
methods, a variant of inexact Newton methods, will be outlined
including two techniques for numerically representing the Jacobian
commonly used in today's applications. Following this, a new algorithm
utilizing Monte Carlo Synthetic Acceleration will be presented and
implementation details will be discussed. Using the new algorithm and
a production fluid physics code, a group of challenging nonlinear
benchmark problems for the Navier-Stokes equations will be used to
verify the correctness of the method relative to solutions obtained
using conventional methods in various flow regimes and
geometries. Finally, we will compare the performance of the new method
against contemporary Newton-Krylov methods for the same benchmark
problems.

%%---------------------------------------------------------------------------%%
\section{Preliminaries\ }
\label{sec:nonlinear_preliminaries}
We formulate the \textit{nonlinear problem} as follows
\cite{knoll_jacobian-free_2004}:
\begin{equation}
  \ve{F}(\ve{u}) = \ve{0}\:,
  \label{eq:nonlinear_problem}
\end{equation}
where $\ve{u} \in \mathbb{R}^n$ is the solution vector and
$\ve{F}:\mathbb{R}^N \rightarrow \mathbb{R}^N$ is the function of
nonlinear residuals. We write the nonlinear system in this form so
that when an exact solution for $\ve{u}$ is achieved, all residuals
evaluate to zero. \textit{Newton's method} is a root finding algorithm
and therefore we can use it to solve Eq~(\ref{eq:nonlinear_problem})
if we interpret the exact solution $\ve{u}$ to be the roots of
$\ve{F}(\ve{u})$. Newton's method is also an iterative scheme, and we
can generate this procedure by building the Taylor expansion of the
$k+1$ iterate, $\ve{F}(\ve{u}^k)$, of the nonlinear residuals about
the previous $k$ iterate, $\ve{F}(\ve{u}^{k+1})$:
\begin{equation}
  \ve{F}(\ve{u}^{k+1}) = \ve{F}(\ve{u}^{k}) +
  \ve{F}'(\ve{u}^{k})(\ve{u}^{k+1}-\ve{u}^{k}) +
  \frac{\ve{F}''(\ve{u}^{k})}{2}(\ve{u}^{k+1}-\ve{u}^{k})^2 + \cdots
  \:.
  \label{eq:newton_derivation_1}
\end{equation}
If we ignore the nonlinear terms in the expansion and assert that at
the $k+1$ iterate $\ve{u}^{k+1}$ is the exact solution such that
$\ve{F}(\ve{u}^{k+1}) = \ve{0}$, then we are left with the following
equality:
\begin{equation}
  -\ve{F}(\ve{u}^{k}) =
  \ve{F}'(\ve{u}^{k})(\ve{u}^{k+1}-\ve{u}^{k})\:.
  \label{eq:newton_derivation_2}
\end{equation}
We note two things of importance in
Eq~(\ref{eq:newton_derivation_2}). The first is that
$\ve{F}'(\ve{u}^{k})$ is in fact the \textit{Jacobian},
$\ve{J}(\ve{u})$, of the set of nonlinear residuals and is defined
element-wise as:
\begin{equation}
  J_{ij} = \frac{\partial F_i(\ve{u})}{\partial u_j}\:.
  \label{eq:jacobian_def}
\end{equation}
Second, we note that $(\ve{u}^{k+1}-\ve{u}^{k})$ is simply the
solution update from the $k$ iterate to the $k+1$ iterate. We will
define this update as the \textit{Newton correction} at the $k$
iterate, $\delta \ve{u}^k$. To finish, we can then rearrange
Eq~(\ref{eq:newton_derivation_2}) to define the Newton iteration
scheme for nonlinear problems:
\begin{subequations}
  \begin{gather}
    \ve{J}(\ve{u}) \delta \ve{u}^k = -\ve{F}(\ve{u}^{k})\\
    \ve{u}^{k+1} = \ve{u}^k + \delta \ve{u}^k\:.
  \end{gather}
  \label{eq:newton_iteration}
\end{subequations}

There are then three distinct steps to perform: evaluation of the
nonlinear residuals using the solution at the $k$ iterate, the
solution of a linear system to compute the Newton correction where the
Jacobian matrix of the nonlinear equation set is the linear operator,
and the application of the correction to the previous iterate's
solution to arrive at the next iterate's solution. In the asymptotic
limit, the iterations of Newton's method will converge the nonlinear
residual quadratically \cite{kelley_iterative_1995}. Convergence
criteria is set for stopping the iteration sequence based on the
nonlinear residual. Commonly, the following criteria is used:
\begin{equation}
  ||\ve{F}(\ve{u}^{k})|| < \epsilon ||\ve{F}(\ve{u}^{0})||\:,
  \label{eq:newton_stopping_criteria}
\end{equation}
where $\epsilon$ is a user defined tolerance parameter.

Newton's method is \textit{consistent} in that all components of the
nonlinear functions that describe the physics we are modeling are
updated simultaneously in the iteration sequence with respect to one
another. This is in comparison to \textit{inconsistent} strategies,
such as a pressure correction strategy for solving the Navier-Stokes
equations \cite{pletcher_computational_1997}, where the components of
$\ve{u}$ are updated in a staggered fashion depending on the
particular equations that they are associated with.

\subsection{Inexact Newton Methods}
\label{subsec:inexact_newton_methods}
Inexact Newton methods arise when the Jacobian operator is not exactly
inverted, resulting in an inexact Newton correction as initially
described by Dembo and others \cite{dembo_inexact_1982}. For common
sparse nonlinear systems, which in turn yield a sparse Jacobian
matrix, this situation occurs when conventional iterative methods are
applied. In their definition, Dembo formulated inexact methods such
that they are independent of the linear method used to solve for the
Newton correction and therefore are amenable to use with any linear
solver. Furthermore, they bind the convergence of the outer nonlinear
iteration to the inner linear iteration such that:
\begin{equation}
  ||\ve{J}(\ve{u}^k)\delta \ve{u}^k + \ve{F}(\ve{u}^k)|| \leq \eta^k
  ||\ve{F}(\ve{u}^k)||\:,
  \label{eq:inexact_newton_forcing}
\end{equation}
where $\eta^k \in [0,1)$ is defined as the \textit{forcing term} at
  the $k$ iterate. Eq~(\ref{eq:inexact_newton_forcing}) then states
  that the residual generated by the linear solver is bound by the
  nonlinear residual and how tightly it is bound is defined by the
  forcing term. This is useful in that we can vary how tightly coupled
  the convergence of the linear iterations used to generate the Newton
  correction is to the nonlinear iteration by relaxing or tightening
  the convergence properties on the linear iterative method. As a
  result, strategies for determining the forcing term can vary
  depending on the problem type and can greatly affect the convergence
  of the method or even prohibit convergence
  \cite{eisenstat_choosing_1996}. In addition, \textit{globalization
    methods} may be used to modify the Newton correction in a more
  desirable direction such that convergence properties can be
  improved when the initial guess for $\ve{u}$ is poor
  \cite{pawlowski_globalization_2006}.

\subsection{Newton-Krylov Methods\ }
\label{subsec:newton_krylov_methods}
A form of inexact Newton methods, \textit{Newton-Krylov methods} are
nonlinear iterative methods that leverage a Krylov subspace method as
the linear solver for generating the Newton correction
\cite{kelley_iterative_1995}. As outlined in
Appendix~\ref{ch:linear_problem}, Krylov methods are robust and enjoy
efficient parallel implementations on modern
architectures. Furthermore, their lack of explicit dependence on the
operator make them easier to implement than other
methods. Additionally, although many iterations can become memory
intensive due to the need to store the Krylov subspace for the
orthogonalization procedure, at each nonlinear iteration this cost is
reset as the Jacobian matrix will change due to its dependence on the
solution vector. This means that for every nonlinear iteration, a
completely new linear system is formed for generating the Newton
correction and we can modify the Krylov solver parameters accordingly
to accommodate this. In most nonlinear problems, the Jacobian operator
is generally non-symmetric and therefore either Krylov methods with
long recurrence relations that can handle non-symmetric systems must
be considered or the Newton correction system must be preconditioned
such that the operator is symmetric and short recurrence relation
methods can be potentially be used.

With many Krylov methods available, which to use with the Newton
method is dependent on many factors including convergence rates and
memory usage. Several studies have been performed to investigate this
\cite{mchugh_inexact_1993,knoll_newton-krylov_1995}. In their
numerical studies in 1995, Knoll and McHugh used the set of highly
nonlinear and stiff convection-diffusion-reaction equations to solve a
set of tokamak plasma problems with the goal of measuring solver
performance with Newton's method. They note several trade-offs in
using Krylov methods with the Newton solver. The first is that the
optimization condition that results from the constraints (e.g. the
minimization of the GMRES residual over the Krylov space) can be
relaxed by restricting the size of the subspace such that only a fixed
number of subspace vectors may be maintained, thus reducing memory
requirements.

We can also relax the optimization condition by instead restarting the
recurrence relation with a new set of vectors once a certain number of
vectors have been generated. The optimization condition is maintained
over that particular set of vectors, however, Knoll and McHugh note
that this ultimately slows the convergence rate as compared to keeping
all vectors as the new set of vectors is not necessarily orthogonal to
the previous set, and therefore not optimal over the entire iteration
procedure. The orthogonality condition can be relaxed by using a
recurrence relation that does not generate a strictly orthonormal
basis for the Krylov subspace such as the Lanzcos biorthogonalization
procedure, resulting in memory savings due to the shorter Lanzcos
recurrence relation. Based on their numerical analysis, they observed
GMRES to be the most robust for any given nonlinear problem and
therefore used it for subsequent work in this area.

\subsection{Jacobian-Free Approximation}
\label{subsec:jacobian_free_approximation}
In many cases, the Jacobian is difficult to form from the difference
equations and costly to evaluate for large equation sets. For simple
nonlinear cases such as the Navier-Stokes equations, the derivatives
can be computed and coded, but due to the complexity of those
derivatives and the resulting difference equations this task can be
tedious, error prone, and must be repeated for every equation
set. Furthermore, in their 1995 work, Knoll and McHugh also noted that
a dominating part of their computation time was the evaluation of the
difference equations for building the Jacobian
\cite{knoll_newton-krylov_1995}. By recognizing that Krylov methods
only need the action of the operator on a vector instead of the
operator itself, the Jacobian can instead be approximated through
various numerical methods including a difference-based Jacobian-free
formulation.

Jacobian-Free methods, and in particular \textit{Jacobian-Free
  Newton-Krylov} (JFNK) methods \cite{knoll_jacobian-free_2004}, rely
on forming the action of the Jacobian on a vector as required by the
Krylov solver through a forward difference scheme. In this case, the
action of the Jacobian on some vector $\ve{v}$ is given as:
\begin{equation}
  \ve{J}(\ve{u})\ve{v} = \frac{\ve{F}(\ve{u} + \epsilon \ve{v}) -
    \ve{F}(\ve{u})}{\epsilon}\:,
  \label{eq:jacobian_free_product}
\end{equation}
where $\epsilon$ is a small number typically on the order of machine
precision. Kelley \cite{kelley_iterative_1995} points out a potential
downfall of this formulation in that if the discretization error in
$\ve{F}(\ve{u})$ is on the order of the perturbation parameter
$\epsilon$, then the finite difference error from
Eq~(\ref{eq:jacobian_free_product}) pollutes the solution.

In addition, Knoll and McHugh noted that for preconditioning purposes,
part of the Jacobian must still explicitly be formed periodically and
that linear solver robustness issues were magnified by the matrix-free
approach due to the first-order approximation. This formation
frequency coupled with the numerous evaluations of the Jacobian
approximation create a situation where after so many nonlinear
iterations, it becomes cheaper to instead fully form the Jacobian. For
simple equation sets, this may only take 5-10 Newton iterations to
reach this point while over 30 may be required for larger equations
sets and therefore larger Jacobians.

\subsection{Automatic Differentiation for Jacobian Generation}
\label{subsec:automatic_differentiation}
If it is acceptable to store the actual Jacobian matrix, other methods
are available to construct it without requiring hand-coding and
evaluating derivatives, thus eliminating the associated issues. In
addition, if any additional equations are added to the system or a
higher order functional approximation is desired, it would be useful
to avoid regenerating and coding these derivatives. Becoming more
prominent in the 1990's, \textit{automatic differentiation} is a
mechanism by which the derivatives of a function can be generated
automatically by evaluating it. Automatic differentiation is built on
the concept that all functions discretely represented in a computer
are ultimately represented by elementary mathematical operations. If
the chain rule is applied to those elementary operations, then the
derivatives of those functions can be computed to the order of
accuracy of their original discretization in a completely automated
way \cite{averick_computing_1994}.

The work of Bartlett and others \cite{bartlett_automatic_2006}
extended initial Fortran-based work in the area of automatic
differentiation implementations to leverage the parametric type and
operator overloading features of C++ \cite{stroustrup_c++_1997}. They
formulate the differentiation problem from an element viewpoint by
assuming that a global Jacobian can be assembled from local element
function evaluations of $e_k : \mathbb{R}^{n_k} \rightarrow
\mathbb{R}^{m_k}$, similar to the finite element assembly procedure
as:
\begin{equation}
  \ve{J}(\ve{u}) = \sum_{i=1}^N \ve{Q}^T_i \ve{J}_k \ve{P}_i\:,
  \label{eq:fad_global_jacobian}
\end{equation}
where $\ve{J}_{k_i} = \partial e_{k_i} / \partial P_i u$ is the
$k^{th}$ element function Jacobian, $\ve{Q} \in \mathbb{R}^{n_{k_i}
  \times N}$ is a projector onto the element domain and $\ve{P} \in
\mathbb{R}^{m_{k_i} \times N}$ a projector onto the element range for
$\ve{F}(\ve{u}) \in \mathbb{R}^{N \times N}$. The Jacobian matrix for
each element will therefore have entirely local data in a dense
structure, eliminating the need for parallel communication and sparse
techniques during differentiation. Only when all local differentials
are computed does communication of the Jacobian occur through
gather/scatter operations in order to properly assemble it.

Also of benefit is the fact that element-level computations generally
consist of a smaller number of degrees of freedom, thus reducing
memory requirements during evaluation as compared to a global
formulation of the problem. Such a formulation is not limited to
finite element formulations and is amenable to any scheme where the
system is globally sparse with degrees of freedom coupled to local
domains including finite volume representations. The templating
capabilities of C++ were leveraged with the element-based evaluation
and assembly scheme as in Eq~(\ref{eq:fad_global_jacobian}) by
templating element function evaluation code on the evaluation type. If
these functions are instantiated with standard floating point types
then the residual is returned. If they are instead instantiated with
the operator-overloaded automatic differentiation types, both the
residual and Jacobian are returned.

Of interest to Bartlett, Averick, and the many others that have
researched automatic differentiation are measures of its performance
relative to hand-coded derivatives and capturing the Jacobian matrix
from matrix-free approximations. Given their element-based function
evaluation scheme, Bartlett's work varied the number of degrees of
freedom per element and compared both the floating point operation
count and CPU time for both the templated automatic differentiation
method and hand-coded derivatives for Jacobian evaluations. Although
they observed a 50\% increase in floating point operations in the
templated method over the hand-coded method, run times were observed
to be over 3 times faster for the templated method due to the fact
that the element-based formulation of the templated method is causing
better utilization of cache and therefore faster data
access. Furthermore, they observed linear scaling behavior for
automatic differentiation as the number of degrees of freedom per
element were increased. Based on these results, this type of automatic
differentiation formulation was deemed acceptable for use in
large-scale, production physics codes.

%%---------------------------------------------------------------------------%%
\section{The FANM Method\ }
\label{sec:fanm}
In physics codes based on nonlinear equations sets, Newton-Krylov
methods are the primary means of generating a fully consistent
solution scheme
\cite{evans_development_2006,evans_enhanced_2007,gaston_parallel_2009,godoy_parallel_2012}. It
is common that for large scale simulations these problems are memory
limited due to the subspaces generated by robust Krylov methods which
may often build hundreds of vectors during a Newton step. Often, a
matrix-free approach is chosen to relax memory requirements over
directly generating the Jacobian matrix and facilitate the
implementation. However, as we observed in previous sections, these
matrix-free methods suffer from poorly scaled problems and the first
order error introduced by the Jacobian approximation. In addition, it
was observed that the savings induced by the matrix-free approach is
eventually amortized over a number of nonlinear iterations where it
becomes more efficient computationally to instead form the Jacobian.

In Chapter~\ref{ch:stochastic_methods}, we focused our efforts on
developing and improving Monte Carlo methods for inverting linear
systems. These methods, when used to accelerate a stationary method in
MCSA, enjoy exponential convergence rates. Although this requires more
storage to represent the linear system than that of a Krylov method
where the operator is not required, we do not incur any additional
storage costs once the iteration sequence begins. In the context of
nonlinear problems, the Jacobian matrix that we are required to
generate for the Monte Carlo solvers may be generated at will from the
nonlinear functions in the Newton system using automatic
differentiation. Not only do we then have a simple and automated way
to generate the Jacobian, but we also enjoy a Jacobian of numerical
precision equivalent to that of our function evaluations.

We therefore propose the \textit{Forward-Automated Newton-MCSA} (FANM)
method that utilizes all of the above components. Presented in
Algorithm~\ref{alg:fanm}, the FANM method is an inexact Newton method
where MCSA is used to compute the Newton correction. In line 3,
automatic differentiation is used at each iteration to build the
Jacobian operator which can in turn be used to build weights and
probabilities for the Monte Carlo game. In line 4, MCSA is used to
solve the linear problem for the Newton correction. As with other
inexact methods, any given forcing term can be used to control the
convergence of the MCSA iteration at every Newton step. Finally, in
line 5 the Newton correction is applied and the iteration proceeds
until convergence. In addition to forcing term adjustments, other
techniques for improving performance of the nonlinear iteration may be
used with FANM including globalization methods.

\begin{algorithm}[h!]
  \caption{FANM Algorithm}
  \label{alg:fanm}
  \begin{algorithmic}[1]
    \State $k := 0$ 
    \While{$||\ve{F}(\ve{u}^{k})|| > \epsilon
      ||\ve{F}(\ve{u}^{0})||$} 
    \State $\ve{J}(\ve{u}^{k}) \leftarrow AD(\ve{F}(\ve{u}^k))$ 
    \Comment{Automatic differentiation} 
    \State $\ve{J}(\ve{u}^k) \delta \ve{u}^k = -\ve{F}(\ve{u}^{k})$
    \Comment{Solve for the Newton correction with MCSA} 
    \State $\ve{u}^{k+1} \leftarrow \ve{u}^k + \delta \ve{u}^k$ 
    \State $k \leftarrow k+1$ 
    \EndWhile
  \end{algorithmic}
\end{algorithm}

At first glance, FANM is simply a variant of an inexact Newton method
with a specific requirement of a fully formed Jacobian. However, FANM
has the potential to provide value in several areas. First, if
iterative or parallel performance of MCSA can be demonstrated to
better than that of GMRES or similar subspace methods for asymmetric
systems, FANM may outperform a Newton-Krylov method. Second, for
situations where memory is a constraint, especially in situations
where the Jacobian is generated often if not at every Newton
iteration, not building a subspace for ill-conditioned problems will
alleviate the associated cost. However, until the preconditioning
issues associated with MCSA as presented in the previous chapter using
the $SP_N$ equations can be resolved, this hypothesis can
unfortunately not be researched. Explicit algebraic preconditioning
used in conventional solution methods will almost certainly be
required to reduce the spectral radius of most Jacobian operators for
MCSA to apply and therefore the current preconditioning strategy will
destroy all potential memory benefits of FANM.

%%---------------------------------------------------------------------------%%
\section{Navier-Stokes Benchmark Problems\ }
\label{sec:ns_benchmarks}
To verify the FANM method for nonlinear problems, we choose benchmark
solutions for the 2-dimensional, steady, incompressible Navier-Stokes
equations on a rectilinear grid in much the same way as Shadid and
Pawlowski's work on Newton-Krylov methods for the solution of these
equations \cite{shadid_inexact_1997,pawlowski_globalization_2006}. We
define these equations as follows:
\begin{subequations}
  \begin{gather}
    \rho \ve{u} \cdot \nabla \ve{u} - \nabla \cdot \ve{T} - \rho
    \ve{g} = \ve{0}
    \label{eq:ns_momentum}\\
    \nabla \cdot \ve{u} = 0
    \label{eq:ns_continuity}\\
    \rho C_p \ve{u} \cdot \nabla T + \nabla \cdot \ve{q} = 0\:,
    \label{eq:ns_energy}
  \end{gather}
  \label{eq:navier_stokes}
\end{subequations}
where $\rho$ is the fluid density, $\ve{u}$ is the fluid velocity,
$C_p$ the specific heat capacity at constant pressure of the fluid,
$T$ the temperature of the fluid, and $\ve{g}$ the acceleration due to
gravity. Eq~(\ref{eq:ns_momentum}) provides momentum transport,
Eq~(\ref{eq:ns_continuity}) provides the mass balance, and
Eq~(\ref{eq:ns_energy}) provides energy transport with viscous
dissipation effects neglected. In addition, we close the system with
the following equations:
\begin{subequations}
  \begin{gather}
    \ve{T} = -P \ve{I} + \mu[\nabla \ve{u} + \nabla \ve{u}^T]
    \label{eq:ns_stress_tensor}\\
    \ve{q} = - k \nabla T\:,
    \label{eq:ns_heat_flux}
  \end{gather}
  \label{eq:ns_closure}
\end{subequations}
where $\ve{T}$ is the stress tensor, $P$ is the hydrodynamic pressure,
$\mu$ is the dynamic viscosity of the fluid, $\ve{q}$ is the heat flux
in the fluid, and $k$ is the thermal conductivity of the fluid. This
set of strongly coupled equations possesses both the nonlinearities
and asymmetries that we are seeking for qualification of the FANM
method. Further, physical parameters within these equations can be
tuned to enhance the nonlinearities. We will then apply these
equations to the following three standard benchmark problems that
cover convection dominated flow and flow dominated by inertial
forces in two different geometries.

\subsection{Thermal Convection Cavity Problem}
\label{subsec:natural_convection_cavity}
In 1983 a set of benchmark solutions for the natural convection of air
in a square cavity was published \cite{de_vahl_davis_natural_1983} as
shown in Figure~\ref{fig:natural_convection_cavity} for the solution
of the energy, mass, and momentum equations.

\begin{figure}[t!]
  \begin{center}
    \scalebox{1.5}{
      \input{chapters/nonlinear_problem/natural_convection_cavity.pdftex_t} }
  \end{center}
  \caption{\textbf{Problem setup for the natural convection cavity
      benchmark.} \textit{Dirichlet conditions are set for the
      temperature on the left and right while Neumann conditions are
      set on the top and bottom of the Cartesian grid. The temperature
      gradients will cause buoyancy-driven flow. Zero velocity
      Dirichlet conditions are set on each boundary. No thermal source
      was present.}}
  \label{fig:natural_convection_cavity}
\end{figure}

In this problem, a rectilinear grid is applied to the unit square. No
heat flow is allowed out of the top and bottom of the square with a
zero Neumann condition specified. Buoyancy driven flow is generated by
the temperature gradient from the hot and cold Dirichlet conditions on
the left and right boundaries of the box. By adjusting the Rayleigh
number of the fluid (and therefore adjusting the ratio of convective
to conductive heat transfer), we can adjust the influence of the
nonlinear convection term in Eq~(\ref{eq:ns_momentum}). In Shadid's
work, Rayleigh numbers of up to \sn{1}{6} were used for this benchmark.

Isotherms of the temperature solution for these equations are given in
Figure~\ref{fig:convection_isotherms} for Rayleigh numbers of
\sn{1}{3}, \sn{1}{4}, \sn{1}{5} and \sn{1}{6}. Note that as the
Rayleigh number is increased, the fluid begins to rotate in the
clockwise direction. This causes the temperature gradients to rotate
as well with the fluid, given the increased rotation observed for the
isotherms in the figure. As the Rayleigh number increases, thermal
energy and momentum transport in the system become more dominated by
convection rather than diffusion, causing the nonlinearities in the
advective derivatives to grow.

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/nonlinear_problem/convection_isotherms.png}
  \end{center}
  \caption{\textbf{Isotherms from solution to the thermal convection
        cavity problem.} \textit{Top left: Ra = \sn{1}{3}; Top right:
        Ra = \sn{1}{4}; Bottom left: Ra = \sn{1}{5}; Bottom left: Ra =
        \sn{1}{6}. Isotherms on unit temperature scale from 0 to 1
        with 10 uniformly spaced curves for each calculation.}}
  \label{fig:convection_isotherms}
\end{figure}

\clearpage

\subsection{Lid Driven Cavity Problem}
\label{subsec:lid_driven_cavity}
To study performance for flow driven by inertial forces, the second
benchmark problem given by Ghia \cite{ghia_high-re_1982} incorporates
a driver for the flow to introduce higher Reynolds numbers into the
system, providing more inertial force to overcome the viscous forces
in the fluid. The setup for this problem is equally simple, containing
only the Dirichlet conditions as given in
Figure~\ref{fig:lid_driven_cavity} and is only applied to the mass and
momentum equations on the unit square.

\begin{figure}[t!]
  \begin{center}
    \scalebox{1.5}{
      \input{chapters/nonlinear_problem/lid_driven_cavity.pdftex_t} }
  \end{center}
  \caption{\textbf{Problem setup for the lid driven cavity benchmark.}
    \textit{Dirichlet conditions of zero are set for the velocity on
      the left and right and bottom while the Dirichlet condition set
      on the top provides a driving force on the fluid.}}
  \label{fig:lid_driven_cavity}
\end{figure}

The top boundary condition will provide a driver for the flow and its
variation will in turn vary the Reynolds number of the fluid. An
increased velocity will generate more inertial forces in the fluid,
which will overcome the viscous forces and again increase the
influence of the nonlinear terms in Eq~(\ref{eq:ns_momentum}). Shadid
used Reynolds numbers up to \sn{1}{4} for this benchmark problem.

Isocurves of the velocity magnitudes are given in
Figure~\ref{fig:driven_velocity_isocurves} for solutions to the lid
driven cavity problem at Reynolds numbers of 100, 300, 500 and
700. Increasing the velocity on the top boundary of the problem drives
up the Reynolds number of the system making it more difficult to
solve. More rotation of the fluid is induced by inertial forces and we
begin to see some additional vortexes form besides the primary vortex
near the center of the box.

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/nonlinear_problem/driven_velocity_isocurves.png}
  \end{center}
  \caption{\textbf{Velocity magnitude isocurves from solution to the
      lid driven cavity problem.} \textit{Top left: Re = 100; Top
      right: Re= 300; Bottom left: Re = 500; Bottom left: Re =
      700. Isocurves on unit velocity scale from 0 to 1 with 20
      uniformly spaced curves for each calculation.}}
  \label{fig:driven_velocity_isocurves}
\end{figure}

\clearpage

\subsubsection{Backward-Facing Step Problem}
\label{subsubsec:backward_facing_step}
The third benchmark was generated by Gartling in 1990 and consists of
both flow over a backward step and an outflow boundary condition
\cite{gartling_test_1990}. Using the mass and momentum equations
while neglecting the energy equation, this problem utilizes a longer
domain with a 1/25 aspect ratio used for this work\footnote{The
  literature on this benchmark suggests using a domain with a 1/30
  aspect ratio to reduce the effect of the outflow boundary condition
  on the upstream recirculation zones. Using a 1/25 aspect ratio for
  this work was not observed to disturb the recirculation zones and
  was therefore used to reduce memory pressure on the FANM solver.}
with the boundary conditions as shown in
Figure~\ref{fig:backward_facing_step}.

\begin{figure}[t!]
  \begin{center}
    \scalebox{1.2}{
      \input{chapters/nonlinear_problem/backward_facing_step.pdftex_t} }
  \end{center}
  \caption{\textbf{Problem setup for the backward facing step
      benchmark.} \textit{Zero velocity boundary conditions are
      applied at the top and bottom of the domain while the outflow
      boundary condition on the right boundary is represented by zero
      stress tensor components in the direction of the flow. For the
      inlet conditions, the left boundary is split such that the top
      half has a fully formed parabolic flow profile and the bottom
      half has a zero velocity condition, simulating flow over a
      step. The $H$ parameter given in the parabolic inlet condition
      is the width of the inlet.}}
  \label{fig:backward_facing_step}
\end{figure}

In this problem, the inflow condition is specified by a fully-formed
parabolic flow profile over a zero velocity boundary representing a
step. For this work, a slightly different parabolic formulation for
the inlet velocities was used than in the benchmark. In the benchmark
solutions, the average velocity in the parabola was 2/3 that of the
maximum velocity while the formula given in
Figure~\ref{fig:backward_facing_step} produces a more plug-like flow
with an average velocity 3/4 that of the maximum velocity. The flow
over this step will generate a recirculating backflow under the inlet
flow towards the step and at higher Reynolds numbers a recirculating
zone will form farther downstream from the step. As in the lid driven
cavity problem, the nonlinear behavior of this benchmark and the
difficulty in obtaining a solution is dictated by the Reynolds number
of the fluid. In Shadid's work, a Reynolds numbers up to 800 were
used.

Compared to the lid driven cavity benchmark, the flow over a backward
step is much more representative of flow driven by inertial forces in
a nuclear reactor that may encounter several obstacles along the flow
path. For example, spacer grids in fuel assemblies or mixing vanes
designed to induce turbulence and improve heat transfer will create
recirculation zones much like those observed for this benchmark
problem. Figure~\ref{fig:step_velocity_isocurves} gives the velocity
magnitude contours for solutions to this problem with Reynolds numbers
of 200, 300, 400, and 500.

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/nonlinear_problem/step_velocity_isocurves.png}
  \end{center}
  \caption{\textbf{Velocity magnitude isocurves from solution to the
      backward facing step problem.} \textit{From top to bottom: Re =
      200, Re = 300, Re = 400, Re = 500. Isocurves on unit velocity
      scale from 0 to 1 with 20 uniformly spaced curves for each
      calculation. The depression forming in the upper right portion
      of visible flow is the second vortex.}}
  \label{fig:step_velocity_isocurves}
\end{figure}

\clearpage

%%---------------------------------------------------------------------------%%
\section{FANM Verification\ }
\label{sec:fanm_verification}

For each of the benchmark problems, we will present the results of
computations using FANM and compare them to results using a
Newton-Krylov method in order to verify the correctness of FANM and
its applicability to fluid problems. Given the difficulty of these
benchmark problems, the linear solve at each Newton step was
preconditioned with an algebraic multigrid method\footnote{The ML
  algebraic multigrid library in the Trilinos scientific computing
  libraries \cite{heroux_overview_2005} was used for all problems.}, a
common technique for preconditioning these equations
\cite{ghia_high-re_1982,evans_enhanced_2007} with the preconditioner
reconstructed at every Newton step for the new Jacobian. Without
multigrid preconditioning, the residual in MCSA was observed to
stagnate. Typically this happens because the Richardson iteration and
Neumann series on which the Monte Carlo method is based will quickly
damp the higher order error modes on the scale of the computational
grid used to solve the problem but will reduce lower order error modes
at a considerably lower rate. In terms of Monte Carlo, this stagnation
manifests itself in histories that simply do no terminate and
therefore no solution is acquired. For many of the problems presented
here, it is likely that a more physics-based preconditioning strategy
such as those presented in
\cite{evans_development_2006,evans_enhanced_2007} will yield better
results for more difficult problems. However, the explicit
preconditioning strategy used here currently limits us to algebraic
preconditioning.

For every benchmark calculation, both FANM and Newton-Krylov were
preconditioned with the same multigrid parameters such that the
conditioning of the system would not be a factor in the verification
of the method or the performance comparison in the following
section. With FANM, this preconditioning was required to reduce the
spectral radius of the Jacobian operators generated at each Newton
step. This preconditioning was also required to obtain good
convergence properties for the Newton-Krylov solutions although it was
not necessary for their convergence. In addition, the same nonlinear
solver parameters including globalization and forcing term strategies
were used with both methods with a nonlinear convergence criteria of a
\sn{1}{-12} norm for the nonlinear residual. As with the radiation
transport calculations in the previous chapter, algebraic
preconditioning with the multigrid method results in dense linear
systems. To mitigate this, the reduced domain approximation was again
applied to the Monte Carlo solution within the MCSA solve occurring at
each nonlinear iteration. The values used for the approximation will
be given for each benchmark problem.

Results in this and the following section were produced by the Drekar
multiphysics code base being developed at Sandia National Laboratories
for fluid flow and other multiple physics simulations
\cite{pawlowski_drekar_2012}. Drekar is a massively parallel physics
framework that utilizes Newton-Krylov methods to generate a fully
consistent solution scheme for the Navier-Stokes equations among other
systems with Jacobian generation through automatic differentiation. To
achieve these results, a Monte Carlo solver generated for this work
was injected into the nonlinear tool chain within Drekar to form an
implementation of the FANM method. For the Newton-Krylov solutions,
the same implementation of GMRES from the Trilinos scientific
computing library Aztec \cite{heroux_overview_2005} used for the
verification of MCSA for the $SP_N$ solutions in the previous chapter
was also used here as a means of comparison and FANM benchmarking.

\subsection{Thermal Convection Cavity Results}
\label{subsec:thermal_convection_verification}

A Drekar model of thermal convection cavity problem was solved using
both the Newton-Krylov solver and the FANM solver for Rayleigh numbers
of \sn{1}{3}, \sn{1}{4}, \sn{1}{5} and \sn{1}{6} on a $50 \times 50$
square grid. Table~\ref{tab:thermal_convection_parameters} gives the
physical parameters of the problem and
Table~\ref{tab:convection_mcsa_parameters} gives the MCSA solver
parameters used within the FANM method. Variation in the Rayleigh
number of the system was achieved by keeping the parameters given in
Table~\ref{tab:thermal_convection_parameters} constant and varying
temperature of the hot side of the cavity. In addition to multigrid
preconditioning for the linear solution at each Newton step,
backtracking as outlined in Shadid's work \cite{shadid_inexact_1997}
was used as a means of globalization to assist the Newton correction
computation while the forcing term at all Rayleigh numbers but
\sn{1}{6} was computed at each Newton step using method 2 from
Shadid's work. In this forcing term method, nonlinear residuals from
the two previous Newton steps are used to inform the choice of the
forcing term for the next step. At a Rayleigh number of \sn{1}{6}, a
constant forcing term of \sn{1}{-4} was use to give better convergence
properties.

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lcc}\hline\hline
      \multicolumn{1}{l}{Parameter}& 
      \multicolumn{1}{c}{Value}&
      \multicolumn{1}{c}{Units}\\\hline
      $x_{min}$ & 0.0 & $m$ \\
      $x_{max}$ & 1.0 & $m$ \\
      $y_{min}$ & 0.0 & $m$ \\
      $y_{max}$ & 1.0 & $m$ \\
      $N_x$ & 50 & - \\
      $N_y$ & 50 & - \\
      $DOFs$ & 10404 & - \\
      Initial Temperature & 0.0 & $K$ \\
      Density & 1.0 & $kg / m^3$ \\
      Specific Heat & 1.0 & J / $(kg \times K)$ \\
      Dynamic Viscosity & 0.71 & $(N \times s) / m^2$ \\
      Thermal Conductivity & 1.0 & $W / (m \times K)$ \\
      Thermal Expansion Coefficient & 1000 & $1 / K$ \\
      Thermal Diffusivity & 1.0 & $m^2 / s$ \\
      Kinematic Viscosity & 0.71 & $m^2 / s$ \\
      Gravitational Acceleration & 10 & $m / s^2$ \\
      Cavity Characteristic Length & 1.0 & $m$ \\
      Prandtl Number & 0.71 & - \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Thermal convection cavity FANM verification model
      problem parameters.}  \textit{The Navier-Stokes equations in 2
      dimensions are used for the model problem. The Rayleigh number
      is varied by modifying the hot boundary temperature to induce
      larger temperature gradients and buoyancy-driven flow.}}
  \label{tab:thermal_convection_parameters}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lc}\hline\hline
      \multicolumn{1}{l}{Parameter}& 
      \multicolumn{1}{c}{Value}\\\hline
      Histories & 500,000 per iteration \\
      Weight Cutoff & \sn{1}{-2} \\
      Fixed Point Iteration & Richardson \\
      Estimator & Adjoint Collision \\
      Reduced Domain Fill Level & 200 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Thermal convection cavity FANM verification MCSA
      solver parameters.} \textit{No relaxation parameters or variance
      reduction techniques were used with MCSA other than the reduced
      domain approximation.}}
  \label{tab:convection_mcsa_parameters}
\end{table}

For the verification calculations, at each Rayleigh number the fluid
velocity in the x and y directions, fluid pressure, and fluid
temperature values calculated Newton-Krylov were compared against the
Newton-Krylov calculations. No variation in solution was noted for
each of the computations larger than the floating point tolerance
reported by Drekar with an absolute difference value of zero observed
at all spatial locations for each quantity.

\subsection{Lid Driven Cavity Results}
\label{subsec:lid_driven_verification}

A Drekar model of thermal convection cavity problem was solved using
both the Newton-Krylov solver and the FANM solver for Reynolds numbers
of 100, 300, 500, and 700 on a $50 \times 50$ square
grid. Table~\ref{tab:lid_driven_parameters} gives the physical
parameters of the problem and Table~\ref{tab:driven_mcsa_parameters}
gives the MCSA solver parameters used within the FANM
method. Variation in the Reynolds number of the system was achieved by
keeping the parameters given in Table~\ref{tab:lid_driven_parameters}
fixed and varying the fluid velocity in the $x$ direction on the top
boundary. In addition to multigrid preconditioning for the linear
solution at each Newton step, a polynomial line search method was used
as a means of globalization, assisting the Newton correction
computation as outlined by Pawlowski in
\cite{pawlowski_globalization_2006} while the forcing term was again
computed using method 2 from Shadid's work \cite{shadid_inexact_1997}
at all Reynolds numbers.

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lcc}\hline\hline
      \multicolumn{1}{l}{Parameter}& 
      \multicolumn{1}{c}{Value}&
      \multicolumn{1}{c}{Units}\\\hline
      $x_{min}$ & 0.0 & $m$ \\
      $x_{max}$ & 1.0 & $m$ \\
      $y_{min}$ & 0.0 & $m$ \\
      $y_{max}$ & 1.0 & $m$ \\
      $N_x$ & 50 & - \\
      $N_y$ & 50 & - \\
      $DOFs$ & 7803 & - \\
      Density & 1.0 & $kg / m^3$ \\
      Dynamic Viscosity & 0.71 & $(N \times s) / m^2$ \\
      Kinematic Viscosity & 0.71 & $m^2 / s$ \\
      Cavity Characteristic Length & 1.0 & $m$ \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Lid driven cavity FANM verification model
      problem parameters.}  \textit{The Navier-Stokes equations in 2
      dimensions are used for the model problem. The Reynolds number
      is varied by modifying the velocity magnitude on the upper
      boundary to induce flow driven by inertial forces.}}
  \label{tab:lid_driven_parameters}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lc}\hline\hline
      \multicolumn{1}{l}{Parameter}& 
      \multicolumn{1}{c}{Value}\\\hline
      Weight Cutoff & \sn{1}{-2} \\
      Fixed Point Iteration & Richardson \\
      Estimator & Adjoint Collision \\
      Histories, Re=100 & 500,000 per iteration \\
      Histories, Re=300 & 500,000 per iteration \\
      Histories, Re=500 & 500,000 per iteration \\
      Histories, Re=700 & 1,000,000 per iteration \\
      Reduced Domain Fill Level, Re=100 & 200 \\
      Reduced Domain Fill Level, Re=300 & 200 \\
      Reduced Domain Fill Level, Re=500 & 200 \\
      Reduced Domain Fill Level, Re=700 & 300 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Lid driven cavity FANM verification MCSA solver
      parameters.} \textit{No relaxation parameters or variance
      reduction techniques were used with MCSA other than the reduced
      domain approximation. Different values of fill level for the
      reduced domain approximation and histories per iteration were
      used at different Reynolds numbers due to convergence
      requirements.}}
  \label{tab:driven_mcsa_parameters}
\end{table}

In the literature on this benchmark, solutions at Reynolds numbers of
up to \sn{1}{4} were achieved using a variety of nonlinear solution
techniques. With the FANM method, at Reynolds numbers above 700
convergence of the linear iterations could not be achieved regardless
of how aggressive the multigrid preconditioner was set. In these cases,
the Jacobian operators are ill-conditioned enough that even GMRES
converges slowly relative to the thermal convection cavity
benchmark. Therefore, the spectral radius limitation on MCSA is
extremely prohibitive in this case and others where transport driven
by inertial forces is dominant. Because of this, only solutions to
the lid driven cavity benchmark at Reynolds numbers of up to 700 are
presented.

For the verification calculations, at each Reynolds number the fluid
velocity in the x and y directions and fluid pressure values
calculated Newton-Krylov were compared against the Newton-Krylov
calculations. No variation in solution was noted for each of the
computations larger than the floating point tolerance reported by
Drekar with an absolute difference value of zero observed at all
spatial locations for each quantity.

\subsection{Backward Facing Step Results}
\label{subsec:backward_step_verification}

A Drekar model of the backward facing step problem was solved using
both the Newton-Krylov solver and the FANM solver for Reynolds numbers
of 200, 300, 400, and 500 on a $20 \times 250$ rectangular
grid. Table~\ref{tab:backward_step_parameters} gives the physical
parameters of the problem and Table~\ref{tab:step_mcsa_parameters}
gives the MCSA solver parameters used within the FANM
method. Variation in the Reynolds number of the system was achieved by
keeping the parameters given in
Table~\ref{tab:backward_step_parameters} fixed and varying the maximum
fluid velocity in the parabolic inlet condition. In addition to
multigrid preconditioning for the linear solution at each Newton step,
a polynomial line search method was used as a means of globalization
while the forcing term was again computed using method 2 from Shadid's
work \cite{shadid_inexact_1997}.

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lcc}\hline\hline
      \multicolumn{1}{l}{Parameter}& 
      \multicolumn{1}{c}{Value}&
      \multicolumn{1}{c}{Units}\\\hline
      $x_{min}$ & 0.0 & $m$ \\
      $x_{max}$ & 25.0 & $m$ \\
      $y_{min}$ & 0.0 & $m$ \\
      $y_{max}$ & 1.0 & $m$ \\
      $N_x$ & 20 & - \\
      $N_y$ & 250 & - \\
      $DOFs$ & 15813 & - \\
      Density & 1.0 & $kg / m^3$ \\
      Dynamic Viscosity & 1.0 & $(N \times s) / m^2$ \\
      Kinematic Viscosity & 1.0 & $m^2 / s$ \\
      Channel Width & 1.0 & $m$ \\
      Step Height & 0.5 & $m$ \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Backward facing step FANM verification model
      problem parameters.}  \textit{The Navier-Stokes equations in 2
      dimensions are used for the model problem. The Reynolds number
      is varied by modifying the maximum velocity magnitude in the
      parabolic inlet condition.}}
  \label{tab:backward_step_parameters}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lc}\hline\hline
      \multicolumn{1}{l}{Parameter}& 
      \multicolumn{1}{c}{Value}\\\hline
      Weight Cutoff & \sn{1}{-2} \\
      Fixed Point Iteration & Richardson \\
      Estimator & Adjoint Collision \\
      Histories, Re=200 & 500,000 per iteration \\
      Histories, Re=300 & 500,000 per iteration \\
      Histories, Re=400 & 500,000 per iteration \\
      Histories, Re=500 & 1,000,000 per iteration \\
      Reduced Domain Fill Level, All Re & 200 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Backward facing step FANM verification MCSA solver
      parameters.} \textit{No relaxation parameters or variance
      reduction techniques were used with MCSA other than the reduced
      domain approximation. The same fill level for the reduced domain
      approximation was used at every Reynolds numbers and different
      histories per iteration were used at different Reynolds numbers
      due to convergence requirements.}}
  \label{tab:step_mcsa_parameters}
\end{table}

In the literature on this benchmark, solutions at Reynolds numbers of
up to 800 were achieved using a variety of nonlinear solution
techniques. In identical fashion to the lid driven cavity problem, at
Reynolds numbers above 500 convergence of the linear iterations could
not be achieved with FANM regardless of how aggressive the multigrid
preconditioner was set. In these cases, the Jacobian operators are
again ill-conditioned enough that a spectral radius of less than one
could not be achieved with multigrid preconditioning and several cycle
applications. Therefore, the spectral radius limitation on MCSA is
extremely prohibitive in this case as well and even more so than the
lid driven cavity problem where solutions at higher Reynolds numbers
were achieved. Because of this, only solutions to the backward facing
step benchmark at Reynolds numbers of up to 500 are presented.

For the verification calculations, at each Reynolds number the fluid
velocity in the x and y directions and fluid pressure values
calculated Newton-Krylov were compared against the Newton-Krylov
calculations. No variation in solution was noted for each of the
computations larger than the floating point tolerance reported by
Drekar with an absolute difference value of zero observed at all
spatial locations for each quantity.

%%---------------------------------------------------------------------------%%
\section{FANM Performance Comparison to Conventional Methods\ }
\label{sec:fanm_comparison}

Using the benchmarks from the verification, we now compare the
performance of the FANM method against a conventional Newton-Krylov
(NK) method. For each benchmark, we will analyze the iterative
performance of both the nonlinear Newton solver and the linear solver
used to compute the Newton correction vector at each step. Limiting
the number of nonlinear iterations required for solution is critical
for efficiency as nonlinear function evaluations, Jacobian generation,
and preconditioner construction are typically required at every Newton
iteration. In addition, timing results will also be discussed. To
quantify the acceleration provided by the residual Monte Carlo
component of MCSA, each calculation was also performed using the same
Newton method but with a plain preconditioned Richardson iteration as
the linear solution technique to give a Newton-Richardson (NR)
scheme. For all cases the Richardson iteration was preconditioned
in the same way as GMRES and MCSA.

\subsection{Thermal Convection Cavity Results}
\label{subsec:thermal_convection_comparison}

At each Rayleigh number used for the thermal convection cavity
problem, Table~\ref{tab:convection_nonlinear_iter_comparison} gives
the number of Newton iterations required for convergence for each
method.  For all Rayleigh numbers except \sn{1}{5} the number of
Newton iterations required to converge is the same while the \sn{1}{5}
result required an extra FANM iteration. This is due to the fact that
the convergence of the linear residual has a strong effect on the
convergence of the nonlinear residual. For nonlinear calculations,
this is much stronger than the effect on the convergence of the
eigenvalue in the $SP_N$ computations where we always observed that
identical numbers of k-eigenvalue iterations were required to converge
each variation of the problem.

As outlined in our discussion on inexact Newton methods, depending on
how converged the linear residual is at each step, it is possible to
throw off the nonlinear iteration and decrease iterative
performance. Although the same preconditioner was used with both the
Newton-Krylov and FANM calculations, the convergence properties of
MCSA and GMRES are different and we therefore expect them to produce
different results at each nonlinear iteration which will in turn
affect the forcing term calculation. If they are different enough, it
is possible that another Newton iteration may be required. Even
considering the extra Newton iteration, the solutions are converged to
effectively the same residual norm tolerance and produce the same
results as provided by the verification.

Figure~\ref{fig:ra1e5_convergence} gives the nonlinear residual
behavior for the \sn{1}{5} Rayleigh number case that required the
extra FANM iteration. We see that the extra iteration comes from
slightly slower quadratic convergence with FANM for this particular
case. The difference in convergence of the nonlinear problem comes
from the fact that the same multigrid preconditioner parameters are
used for this case as for the \sn{1}{3} and \sn{1}{4} Rayleigh number
cases for the linear solutions. Using the same preconditioner for the
harder \sn{1}{5} case means that the problem is still growing in
difficulty even though it is preconditioned in an identical
manner. This means that the GMRES solver at each Newton step is
converging the linear residual to a smaller tolerance than the MCSA
solver as the forcing terms adapt to the convergence of the method
over the course of nonlinear iterations. For this problem, tighter
convergence of the linear model at later Newton steps also improved
convergence of nonlinear model, resulting in the extra FANM iteration.

As the Rayleigh number increases, the problem becomes more difficult
to solve and therefore a more aggressive preconditioner should
maintain iterative performance of the nonlinear method. This was
observed to be true for the \sn{1}{6} Rayleigh number case. For this
calculation, the number of multigrid cycle applications was bumped up
to 3 from 1 and FANM converged in the same number of nonlinear
iterations as the Newton-Krylov method.

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{cccc}\hline\hline
      \multicolumn{1}{c}{Rayleigh Number}& 
      \multicolumn{1}{c}{NK Iterations}&
      \multicolumn{1}{c}{FANM Iterations}&
      \multicolumn{1}{c}{NR Iterations}\\
      \hline
      \sn{1}{3} & 5 & 5 & 5\\
      \sn{1}{4} & 7 & 7 & 7\\
      \sn{1}{5} & 9 & 10 & 9\\
      \sn{1}{6} & 11 & 11 & 11\\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Thermal convection cavity nonlinear iteration
      results for the performance comparison.} \textit{The results
      presented here were obtained from the benchmark verification
      calculations.}}
  \label{tab:convection_nonlinear_iter_comparison}
\end{table}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/nonlinear_problem/ra_1e5_convergence.pdf}
  \end{center}
  \caption{\textbf{Thermal convection cavity nonlinear residual norm
      as a function of Newton iteration for the \sn{1}{5} Rayleigh
      number case.} \textit{The FANM calculation required an
      additional nonlinear iteration to converge.}}
  \label{fig:ra1e5_convergence}
\end{figure}

In addition to nonlinear iterative performance, performance of the
inner linear solves is also of interest. Here, we seek to maintain a
minimum number of total linear solver iterations over the course of
all nonlinear iterations in order to limit matrix-vector multiplies
and other costly
operations. Table~\ref{tab:convection_linear_iter_comparison} gives
the total number of GMRES iterations required to converge the problem
for the Newton-Krylov method and the total number of MCSA iterations
for the FANM method at each Rayleigh number used. For all cases, fewer
MCSA iterations were required than GMRES iterations to converge the
problem even with both solvers preconditioned with the same multigrid
package and the same parameters. For the \sn{1}{6} Rayleigh number
case where a constant forcing term of \sn{1}{-4} was kept, this
difference is most drastic with FANM requiring only 64\% of the linear
iterations required by the Newton-Krylov scheme with the linear solver
per nonlinear iteration given for this case in
Figure~\ref{fig:ra1e6_linear_iters}. Even the \sn{1}{5} Rayleigh
number case that required more Newton iterations required 5 fewer MCSA
iterations to converge than GMRES iterations. In all cases, MCSA
improved on the iterative performance of using only the Richardson
iteration through the residual Monte Carlo acceleration. If the performance of
the MCSA implementation used here were to be potentially optimized in
the future, this improved iterative performance could also mean
improved time to solution.

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{cccc}\hline\hline
      \multicolumn{1}{c}{Rayleigh Number}&
      \multicolumn{1}{c}{GMRES Iterations}&
      \multicolumn{1}{c}{MCSA Iterations}&
      \multicolumn{1}{c}{Richardson Iterations}\\
      \hline
      \sn{1}{3} & 23 & 18 & 38 \\
      \sn{1}{4} & 23 & 17 & 34 \\
      \sn{1}{5} & 25 & 20 & 34 \\
      \sn{1}{6} & 39 & 25 & 48  \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Thermal convection cavity total sum of linear
      iterations for all nonlinear iterations.} \textit{The results
      presented here were obtained from the benchmark verification
      calculations.}}
  \label{tab:convection_linear_iter_comparison}
\end{table}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/nonlinear_problem/convection_ra1e6_iters.pdf}
  \end{center}
  \caption{\textbf{Thermal convection cavity linear solver iterations
      as a function of Newton iteration for the \sn{1}{6} Rayleigh
      number case.} \textit{MCSA shows better iterative performance
      than GMRES at all but the second nonlinear iteration and MCSA
      always accelerated the Richardson iteration. The forcing term
      was fixed at \sn{1}{-4} at all nonlinear iterations.}}
  \label{fig:ra1e6_linear_iters}
\end{figure}

Although the iterative performance of FANM is excellent for the
thermal convection cavity problem, the explicit preconditioning
strategy causes the same issues as was observed for the $SP_N$
equations in the previous
chapter. Table~\ref{tab:convection_speedup_comparison} gives the
speedup of the Newton-Krylov method over FANM. For all Rayleigh
numbers, we see a fairly constant speedup of O(400) when using
Newton-Krylov over FANM meaning that the implementation of the FANM
method used here is significantly slower. As the inexact Newton scheme
in Drekar was modified by simply swapping out the linear solver used
at each Newton step, this increased compute time is coming directly
from using MCSA as the linear solver and the overhead due to
preconditioning. When preconditioning overhead for the $SP_N$
equations was considered as shown by the data in
Figure~\ref{fig:spn_comparison_prec_time}, a similar speedup of
approximately two orders of magnitude was noted between the Krylov
methods and MCSA showing consistency between those results and the
convection cavity benchmark results presented. In addition, the fairly
constant speedup values as a function of the primary problem parameter
show that FANM has effectively the same time complexity as a
Newton-Krylov method and the implementation timings simply differ by a
constant value, also consistent with the $SP_N$ observations.

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{cc}\hline\hline
      \multicolumn{1}{c}{Rayleigh Number}& 
      \multicolumn{1}{c}{Newton-Krylov Speedup}\\
      \hline
      \sn{1}{3} & 338 \\
      \sn{1}{4} & 336 \\
      \sn{1}{5} & 346 \\
      \sn{1}{6} & 465 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Thermal convection cavity Newton-Krylov speedup
      over FANM method.} \textit{Speedup values are rounded to the
      nearest integer. The results presented here were obtained from
      the benchmark verification calculations.}}
  \label{tab:convection_speedup_comparison}
\end{table}

\subsection{Lid Driven Cavity Results}
\label{subsec:lid_driven_comparison}

For the lid driven cavity benchmark, performance results were again
compared between the FANM and Newton-Krylov calculations. In this
case, the nonlinear iterative performance results as given by
Table~\ref{tab:driven_nonlinear_iter_comparison} show that the same
number of Newton iterations were required to converge the 100, 300,
and 500 Reynolds number cases while the FANM calculation with a
Reynolds number of 700 actually converged in 4 fewer iterations. The
nonlinear residual norm as a function of Newton iteration for this
case is given in Figure~\ref{fig:re700_convergence}.

This result can be explained by the fact that as given by
Table~\ref{tab:driven_mcsa_parameters}, although the preconditioner
parameters were the same for both calculations, the MCSA iteration was
pushed harder by doubling the number of histories used for the
computation at each nonlinear iteration. These extra histories
provided better convergence of the linear model at each iteration and
did not result in oversolving, thereby increasing the convergence rate
of the nonlinear model as well. This result is significant in that it
shows increasing the number of histories used for each MCSA iteration
can improve the performance of both linear and nonlinear convergence.

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{cccc}\hline\hline
      \multicolumn{1}{c}{Reynolds Number}& 
      \multicolumn{1}{c}{NK Iterations}&
      \multicolumn{1}{c}{FANM Iterations}&
      \multicolumn{1}{c}{NR Iterations}\\
      \hline
      100 & 6 & 6 & 7\\
      300 & 9 & 9 & 9\\
      500 & 11 & 11 & 11\\
      700 & 14 & 10 & 12\\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Lid driven cavity nonlinear iteration
      results for the performance comparison.} \textit{The results
      presented here were obtained from the benchmark verification
      calculations.}}
  \label{tab:driven_nonlinear_iter_comparison}
\end{table}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/nonlinear_problem/re700_convergence.pdf}
  \end{center}
  \caption{\textbf{Lid driven cavity nonlinear residual norm as a
      function of Newton iteration for the 700 Reynolds number case.}
    \textit{The FANM calculation converged in 4 fewer iterations than
      the Newton-Krylov calculation.}}
  \label{fig:re700_convergence}
\end{figure}

The difficulty of this benchmark problem as compared to the thermal
convection cavity problem is readily apparent in the linear iteration
counts as given by Table~\ref{tab:driven_linear_iter_comparison}. As
the opposite of what was observed for that benchmark, lid driven
cavity results show that the linear models were significantly more
ill-conditioned, requiring many more MCSA and GMRES iterations. For
all cases except the 700 Reynolds number case, many more MCSA
iterations were required to converge with the observed GMRES iteration
count 64\% of that observed for MCSA when looking at the 100 Reynolds
number case. For the 700 Reynolds number case, fewer MCSA iterations
were observed due to the fact that 4 fewer nonlinear iterations were
performed. The average number of linear iterations per nonlinear
iteration was still much larger for the FANM calculations. In all
cases, MCSA improved on the iterative performance of using only the
Richardson iteration through the residual Monte Carlo acceleration
even considering the cases with Reynolds numbers of 100 and 700 where
an extra Newton-Richardson iteration was required to converge the
nonlinear model.

As an additional means of analysis for the linear solver iteration
results, we will look at a case where MCSA performed poorly compared
to GMRES. Consider the linear iterations as a function of nonlinear
iteration given in Figure~\ref{fig:driven_re500_iters} and the forcing
term as a function of nonlinear iteration given in
Figure~\ref{fig:driven_re500_forcing} for the 500 Reynolds number
case. At the later Newton iterations where the region of fast
quadratic convergence is reached, we see that MCSA required many more
iterations to converge than GMRES, accounting for the majority of the
extra iterations. However, we see in the forcing term data that
forcing terms at later Newton iterations for MCSA were much smaller
than those for GMRES. This means that MCSA was converged to a tighter
tolerance at these iterations and therefore required the extra
iterations.

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/nonlinear_problem/driven_re500_iters.pdf}
  \end{center}
  \caption{\textbf{Lid driven cavity linear solver iterations as a
      function of Newton iteration for the 500 Reynolds number case.}
    \textit{More MCSA iterations were required to converge the linear
      models at later Newton iterations due to smaller forcing
      terms.}}
  \label{fig:driven_re500_iters}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/nonlinear_problem/driven_re500_forcing.pdf}
  \end{center}
  \caption{\textbf{Lid driven cavity forcing terms as a function of
      Newton iteration for the 500 Reynolds number case.} \textit{Once
      the region of fast quadratic convergence is reached, MCSA
      forcing terms were much smaller.}}
  \label{fig:driven_re500_forcing}
\end{figure}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{cccc}\hline\hline
      \multicolumn{1}{c}{Reynolds Number}& 
      \multicolumn{1}{c}{GMRES Iterations}&
      \multicolumn{1}{c}{MCSA Iterations}&
      \multicolumn{1}{c}{Richardson Iterations}\\
      \hline
      100 & 27 & 42 & 151\\
      300 & 35 & 52 & 133\\
      500 & 41 & 56 & 154\\
      700 & 21 & 14 & 32\\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Lid driven cavity total sum of linear iterations
      for all nonlinear iterations.} \textit{The results presented
      here were obtained from the benchmark verification
      calculations. MCSA always accelerated the Richardson
      iteration.}}
  \label{tab:driven_linear_iter_comparison}
\end{table}

As a final comparison, the timing of both methods was again compared
with the speedup of using the Newton-Krylov solver over FANM reported
in Table~\ref{tab:driven_speedup_comparison}. The speedup values
reported here are of the same magnitude as those reported for the
thermal convection cavity problem and again in agreement with those
observed for the $SP_N$ calculations. Lack of significant variation in
these timing results also point to a qualitatively similar time
complexity for the FANM method as a function of primary problem
parameter relative to the Newton-Krylov method. In all cases, the
preconditioning strategy must be significantly improved along with
implementation optimization to avoid these large overheads.

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{cc}\hline\hline
      \multicolumn{1}{c}{Reynolds Number}& 
      \multicolumn{1}{c}{Newton-Krylov Speedup}\\
      \hline
      100 & 299 \\
      300 & 322 \\
      500 & 288 \\
      700 & 488 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Lid driven cavity Newton-Krylov speedup over FANM
      method.} \textit{Speedup values are rounded to the nearest
      integer. The results presented here were obtained from the
      benchmark verification calculations.}}
  \label{tab:driven_speedup_comparison}
\end{table}

\clearpage

\subsection{Backward Facing Step Results}
\label{subsec:backward_step_comparison}

The backward step problem exhibited different performance
characteristics than the other two benchmark problems, likely due to
the different geometry and boundary
conditions. Table~\ref{tab:step_nonlinear_iter_comparison} gives the
nonlinear iteration results for this benchmark. At low Reynolds
numbers, FANM converges in fewer nonlinear iterations than the
Newton-Krylov solver. However, as the problem becomes more difficult
to solver at higher rates of flow, the Newton-Krylov solver eventually
converges in fewer nonlinear iterations. We see the same behavior as
well in the linear solver iteration data given by
Table~\ref{tab:step_linear_iter_comparison}. At the 3 lowest Reynolds
numbers, FANM required significantly fewer linear solver iterations to
converge the problem while the ill-conditioned case at a Reynolds
number of 500 is observed to use many more MCSA iterations. For
Reynolds numbers of both 200 and 500, FANM also reduced the number of
Newton steps need when compared to the Newton-Richardson scheme.

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{cccc}\hline\hline
      \multicolumn{1}{c}{Reynolds Number}& 
      \multicolumn{1}{c}{NK Iterations}&
      \multicolumn{1}{c}{FANM Iterations}&
      \multicolumn{1}{c}{NR Iterations}\\
      \hline
      200 & 10 & 9 & 10\\
      300 & 15 & 14 & 15\\
      400 & 10 & 10 & 10\\
      500 & 19 & 20 & 21\\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Backward facing step nonlinear iteration
      results for the performance comparison.} \textit{The results
      presented here were obtained from the benchmark verification
      calculations.}}
  \label{tab:step_nonlinear_iter_comparison}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{cccc}\hline\hline
      \multicolumn{1}{c}{Reynolds Number}& 
      \multicolumn{1}{c}{GMRES Iterations}&
      \multicolumn{1}{c}{MCSA Iterations}&
      \multicolumn{1}{c}{Richardson Iterations}\\
      \hline
      200 & 24 & 13 & 21\\
      300 & 23 & 17 & 21\\
      400 & 18 & 12 & 14\\
      500 & 30 & 52 & 98\\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Backward facing step total sum of linear iterations
      for all nonlinear iterations.} \textit{The results presented
      here were obtained from the benchmark verification
      calculations.MCSA always accelerated the Richardson iteration.}}
  \label{tab:step_linear_iter_comparison}
\end{table}

Unlike the lid driven cavity benchmark, the extra linear solver
iterations are not due to the smaller forcing terms but are rather
truly a function of the ill-conditioning of the Jacobian matrices
produced at each Newton step. We see this by comparing the results at
Reynolds numbers of 400 where FANM outperformed Newton-Krylov and 500
where Newton-Krylov outperformed FANM. Linear iterations as a function
of nonlinear iteration given in Figure~\ref{fig:step_re400_iters} and
the forcing term as a function of nonlinear iteration given in
Figure~\ref{fig:step_re400_forcing} for the 400 Reynolds number case
and the same data is given in Figure~\ref{fig:step_re500_iters} and
Figure~\ref{fig:step_re500_forcing} for the 500 Reynolds number
case. For both cases, FANM always has larger forcing terms and should
therefore use fewer MCSA iterations at each Newton step to converge
the linear model if the iterative performance is comparable. This is
indeed the case for the 400 Reynolds number case where the linear
model is better conditioned and FANM used fewer total linear solver
iterations to converge the nonlinear residual. However, at a Reynolds
number of 500 FANM has larger forcing terms but requires many more
MCSA iterations to converge the last 2 Newton steps. Therefore, these
extra iterations in this case must come from ill-conditioning.

We additionally see the effects of ill-conditioned Jacobians when
looking at the acceleration provided by MCSA over using a
preconditioned Richardson iteration. For the first 3 cases, the
acceleration was not as significant as for the other benchmark
problems. In this case, this means that the Richardson iteration and
its acceleration by a stochastic realization of a Richardson iteration
are not doing the bulk of the work in this problem. Rather, the
algebraic multigrid preconditioning is performing more work at each
iteration and therefore the acceleration is smaller. When the system
becomes even more ill-conditioned in the 500 Reynolds number case
the acceleration has more of an impact.

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/nonlinear_problem/step_re400_iters.pdf}
  \end{center}
  \caption{\textbf{Backward facing step linear solver iterations as a
      function of Newton iteration for the 400 Reynolds number case.}
    \textit{More MCSA iterations were required to converge the linear
      models at later Newton iterations due to ill-conditioning of the
      Jacobian matrices.}}
  \label{fig:step_re400_iters}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/nonlinear_problem/step_re400_forcing.pdf}
  \end{center}
  \caption{\textbf{Backward facing step forcing terms as a function of
      Newton iteration for the 400 Reynolds number case.} \textit{FANM
      forcing terms were always larger.}}
  \label{fig:step_re400_forcing}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/nonlinear_problem/step_re500_iters.pdf}
  \end{center}
  \caption{\textbf{Backward facing step linear solver iterations as a
      function of Newton iteration for the 500 Reynolds number case.}
    \textit{More MCSA iterations were required to converge the linear
      models at later Newton iterations due to ill-conditioning of the
      Jacobian matrices.}}
  \label{fig:step_re500_iters}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/nonlinear_problem/step_re500_forcing.pdf}
  \end{center}
  \caption{\textbf{Backward facing step forcing terms as a function of
      Newton iteration for the 500 Reynolds number case.} \textit{More
      MCSA iterations were required to converge the linear models at
      later Newton iterations due to ill-conditioning of the Jacobian
      matrices.}}
  \label{fig:step_re500_forcing}
\end{figure}

The ill-conditioning of the 500 Reynolds number case is also apparent
in the timing data given in
Table~\ref{tab:step_speedup_comparison}. For the first two cases, we
see qualitatively similar speedup values for the Newton-Krylov solver
relative to the observations for the first two benchmark problems of
O(100). However, the we expect the ill-conditioning to cause the Monte
Carlo histories in the MCSA solver to have longer random walk lengths
based on our MCSA breakdown analysis in the previous chapter. In
addition, we observed that the linear model required many more MCSA
iterations to converge. These longer random walks and more iterations
manifest themselves in the longer CPU time values and therefore the
larger Newton Krylov speedup value of O(1,000) for the 500 Reynolds
number case.

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{cc}\hline\hline
      \multicolumn{1}{c}{Reynolds Number}& 
      \multicolumn{1}{c}{Newton-Krylov Speedup}\\
      \hline
      200 & 400 \\
      300 & 593 \\
      400 & 825 \\
      500 & 1057 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Backward facing step Newton-Krylov speedup
      over FANM method.} \textit{Speedup values are rounded to the
      nearest integer. The results presented here were obtained from
      the benchmark verification calculations.}}
  \label{tab:step_speedup_comparison}
\end{table}

\subsection{Overall Performance Comparison}
\label{subsec:nonlinear_overall_comparison}

As a final overall comparison, we tabulate the results from the
benchmark calculations over the three performance metrics: nonlinear
iterations, linear iterations, and CPU
time. Tables~\ref{tab:benchmark_nonlinear_comparison},
\ref{tab:benchmark_linear_comparison}, and
\ref{tab:benchmark_time_comparison} give the results. In these tables,
the calculations that produced the best performing results are
tallied. In the case of a tie, both are tallied. Looking at the
tabulated results, we see marginally better performance for FANM in
terms of nonlinear iterations with 1 extra case outperforming the
Newton-Krylov method. As observed for each benchmark, the timing
results were always in favor of the Newton-Krylov method due to the
effects of the explicit preconditioning strategy.

For the total linear solver iterations given in
Table~\ref{tab:benchmark_linear_comparison}, the difference is more
dramatic with FANM converging the nonlinear residual in fewer linear
solver iterations for twice as many cases as the Newton-Krylov method
keeping in mind that for all cases both GMRES and MCSA were
preconditioned with identical multigrid parameters using the same
multigrid preconditioning library. This shows that when properly
preconditioned, we can expect performance improvements in terms of
total linear solver iterations required to converge the nonlinear
residual. With future optimization of MCSA implementations, this may
also mean improved time to solution for FANM when compared to
Newton-Krylov methods.

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lcc}\hline\hline
      \multicolumn{1}{l}{Benchmark}& 
      \multicolumn{1}{c}{Newton-Krylov}&
      \multicolumn{1}{c}{FANM}\\
      \hline
      Convection, Ra=\sn{1}{3} & $\times$ & $\times$ \\
      Convection, Ra=\sn{1}{4} & $\times$ & $\times$ \\
      Convection, Ra=\sn{1}{5} & $\times$ & \\
      Convection, Ra=\sn{1}{6} & $\times$ & $\times$ \\
      Lid Driven, Re=100 & $\times$ & $\times$ \\
      Lid Driven, Re=300 & $\times$ & $\times$ \\
      Lid Driven, Re=500 & $\times$ & $\times$ \\
      Lid Driven, Re=700 & & $\times$ \\
      Backward Step, Re=200 & & $\times$ \\
      Backward Step, Re=300 & & $\times$ \\
      Backward Step, Re=400 & $\times$ & $\times$ \\
      Backward Step, Re=500 & $\times$ & \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Navier-Stokes benchmark comparison for nonlinear
      iterations.} \textit{Over all benchmarks, FANM performed better
      in terms of nonlinear iterations for 1 more case than the
      Newton-Krylov method.}}
  \label{tab:benchmark_nonlinear_comparison}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lcc}\hline\hline
      \multicolumn{1}{l}{Benchmark}& 
      \multicolumn{1}{c}{Newton-Krylov}&
      \multicolumn{1}{c}{FANM}\\
      \hline
      Convection, Ra=\sn{1}{3} & & $\times$ \\
      Convection, Ra=\sn{1}{4} & & $\times$ \\
      Convection, Ra=\sn{1}{5} & & $\times$ \\
      Convection, Ra=\sn{1}{6} & & $\times$ \\
      Lid Driven, Re=100 & $\times$ & \\
      Lid Driven, Re=300 & $\times$ & \\
      Lid Driven, Re=500 & $\times$ & \\
      Lid Driven, Re=700 & & $\times$ \\
      Backward Step, Re=200 & & $\times$ \\
      Backward Step, Re=300 & & $\times$ \\
      Backward Step, Re=400 & & $\times$ \\
      Backward Step, Re=500 & $\times$ & \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Navier-Stokes benchmark comparison for total linear
      solver iterations.} \textit{Over all benchmarks, FANM performed
      better in terms of linear solver iterations for twice as many
      cases as the Newton-Krylov method.}}
  \label{tab:benchmark_linear_comparison}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lcc}\hline\hline
      \multicolumn{1}{l}{Benchmark}& 
      \multicolumn{1}{c}{Newton-Krylov}&
      \multicolumn{1}{c}{FANM}\\
      \hline
      Convection, Ra=\sn{1}{3} & $\times$ & \\
      Convection, Ra=\sn{1}{4} & $\times$ & \\
      Convection, Ra=\sn{1}{5} & $\times$ & \\
      Convection, Ra=\sn{1}{6} & $\times$ & \\
      Lid Driven, Re=100 & $\times$ & \\
      Lid Driven, Re=300 & $\times$ & \\
      Lid Driven, Re=500 & $\times$ & \\
      Lid Driven, Re=700 & $\times$ & \\
      Backward Step, Re=200 & $\times$ & \\
      Backward Step, Re=300 & $\times$ & \\
      Backward Step, Re=400 & $\times$ & \\
      Backward Step, Re=500 & $\times$ & \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Navier-Stokes benchmark comparison for CPU time.}
    \textit{For all benchmarks the explicit MCSA preconditioning
      strategy caused significantly larger CPU times for FANM when
      compared to the Newton-Krylov solutions.}}
  \label{tab:benchmark_time_comparison}
\end{table}

\clearpage

%%---------------------------------------------------------------------------%%
\section{Summary\ }
\label{sec:nonlinear_summary}

In this chapter we have developed and explored Monte Carlo Synthetic
Acceleration methods in the context of solutions to the Navier-Stokes
equations. The following are the significant observations and findings.

\begin{itemize}
\item Forward-Automated Newton-MCSA (FANM), a new inexact Newton
  method leveraging Monte Carlo Synthetic Acceleration, has been
  developed
\item The FANM method has been incorporated into the Drekar
  multiphysics production code base developed at Sandia National
  Laboratories
\item The FANM method has been verified to produce the same solutions
  as a production Newton-Krylov method for three difficult benchmark
  problems for the Navier-Stokes equations in different flow regimes
  and geometries
\item The FANM method has better iterative performance than the
  Newton-Krylov method for convection dominated problems, converging
  in fewer linear solver iterations with the same preconditioning for
  high and low Rayleigh numbers
\item The Newton-Krylov method has better iterative performance than
  the FANM method for flow dominated by inertial forces, converging in
  fewer linear solver iterations with the same preconditioning for
  high Reynolds numbers
\item The spectral radius convergence restriction on MCSA was observed
  to be a significant hindrance by preventing solutions to forced flow
  problems at high Reynolds numbers
\item Explicit algebraic preconditioning used to overcome the spectral
  radius restriction of MCSA creates FANM run times
  $O(100)$-$O(1,000)$ slower than the Newton-Krylov solver
\item More Monte Carlo histories at every FANM iteration can reduce
  the number of linear and nonlinear iterations required to converge
  the problem
\item With future optimization and improved preconditioning, reactor
  accident analyses that rely on convection dominated transport or
  steady state calculations requiring driven flow could experience
  improved time to solution with FANM given the improved iterative
  performance observed for problems of these types
\end{itemize}