### 2013-08-23 doc/dissertation/document/backmatter/extra_scaling.tex
moving subomain nuemann ulam results for multiple sets into an appendix

\chapter{Subdomain Neumann-Ulam Scaling with Multiple Sets}
\label{chap:extra_scaling}
As an additional analysis of MCSA using subdomain Monte Carlo, the multiple
sets scaling studies are repeated to determine if any performance
benefits are made from replication by eliminating Monte Carlo
communication. Figure~\ref{fig:titan_strong_subdomain_ms_time} gives
the runtimes for the strong scaling calculation and
Figure~\ref{fig:titan_strong_subdomain_ms} the computed absolute
efficiencies. For the weak scaling studies,
Figure~\ref{fig:titan_weak_subdomain_ms_time} gives the wall times and
Figure~\ref{fig:titan_weak_subdomain_ms} the efficiencies. For both
scaling studies, the results are effectively the same as for the case
with Monte Carlo communication. This is again due to the fact that
adding sets to the problem is not only an exercise in strong scaling
but it also adds additional parallel overhead in the form of the tally
vector reduction across blocks.

From a parallel performance perspective, MCSA may then be best used in
a mode where the Monte Carlo is performed over only the subdomains
with no replication or overlap added to the system. In this mode,
iterative performance is maintained relative to the solutions with
communication even for more ill-conditioned problems with larger
spectral radii. Furthermore, no overlap was required to maintain
iterative performance be letting histories near the boundary take a
few more steps in their random walks. If in the future there are
ill-conditioned problems where using subdomain Monte Carlo reduces the
algorithmic efficiency of MCSA due to the information lost in the
correction vector, overlap would be a viable mechanism to potentially
maintain that iterative performance, keeping in mind the potential
impacts on scalability that were observed in this work. However,
although eliminating all Monte Carlo communication was observed to be
the best mode of use for MCSA, the parallel Monte Carlo methods
developed for this work still provide value for cases where only a
Monte Carlo solution is desired. In those cases, full transport of the
Monte Carlo histories are required in order to generate the correct
solution.

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_strong_subdomain_ms_time.pdf}
  \end{center}
  \caption{\textbf{Strong scaling total wall time with multiple sets
      and subdomain Monte Carlo.} \textit{Adding multiple sets has in
      general a negative effect on the wall time as was observed for
      the case with full Monte Carlo transport.}}
  \label{fig:titan_strong_subdomain_ms_time}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_strong_subdomain_ms.pdf}
  \end{center}
  \caption{\textbf{Strong scaling absolute efficiency for with
      multiple sets and subdomain Monte Carlo.} \textit{Increased wall
      times from adding parallel reduction operations for multiple
      sets reduces parallel scalability at lower core counts but gives
      a higher efficiency at the strong scaling wall.}}
  \label{fig:titan_strong_subdomain_ms}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_subdomain_ms_time.pdf}
  \end{center}
  \caption{\textbf{Weak scaling total wall time with multiple sets and
      subdomain Monte Carlo.} \textit{The addition of the set tally
      reduction is significant with cost of the reduction observed as
      core count grows.}}
  \label{fig:titan_weak_subdomain_ms_time}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_subdomain_ms.pdf}
  \end{center}
  \caption{\textbf{Weak scaling absolute efficiency for with multiple
      sets and subdomain Monte Carlo.} \textit{Weak scaling
      efficiencies are not improved at large core counts when multiple
      sets are added.}}
  \label{fig:titan_weak_subdomain_ms}
\end{figure}