### 2013-08-23 doc/dissertation/document/frontmatter/acks.tex
updating acks

Great thanks are owed to my advisor and friend Paul Wilson for
graciously giving me years of guidance and the opportunity to develop
this work. He provided me with both the resources and challenges to
grow as a researcher. Without his influence and support, this work
would have never been possible.

Tom Evans is responsible for presenting me with the seeds of this work
and for that I thank him. His mentoring over the past few years has
been invaluable and the skills and lessons I have learned from him
have proven indispensable.

Roger Pawlowski and other members of the Consortium for Advanced
Simulation of Light Water Reactors and Trilinos teams as well as many
staff members at Oak Ridge National Laboratory have provided
tremendous resources for my professional and technical development and
have greatly facilitated this work.

This work was performed under appointment to the Nuclear Regulatory
Commission Fellowship program at the University of Wisconsin - Madison
Department of Engineering Physics.

This research used resources of the Oak Ridge Leadership Computing
Facility at the Oak Ridge National Laboratory, which is supported by
the Office of Science of the U.S. Department of Energy under Contract
No. DE-AC05-00OR22725.