### 2013-08-26 doc/dissertation/document/backmatter/estimator_variance.tex
moving estimator variance into an appendix

\chapter{Monte Carlo Estimator Variance\ }
\label{chap:estimator_variance}

In this appendix we briefly discuss the variance of the three primary
Monte Carlo estimators used with the Neumann-Ulam method in this work.

\section{Forward Estimator Variance\ }
\label{subsec:forward_variance}
We can compute the variance of the forward estimator through traditional
methods by defining the variance, $\sigma_i$, for each component in
the solution:
\begin{equation}
  {\sigma_i}^2 = E\{X_i - (\ve{A}^{-1}\ve{b})_i\}^2 = E\{X_i^2\} - x_i^2\:,
  \label{eq:direct_variance_1}
\end{equation}
where the vector exponentials are computed element-wise. Inserting
Eq~(\ref{eq:direct_expectation_value}) gives:
\begin{equation}
  \sigma_i^2 = \sum_{\nu} P_{\nu} X_{i}(\nu)^2 - x_i^2\:,
  \label{eq:direct_variance_2}
\end{equation}
and applying Eq~(\ref{eq:direct_permutation_contribution}) and
expanding the square of the summation:
\begin{equation}
  \sigma_i^2 = \sum_{\nu} P_{\nu} \Big(\sum_{m=0}^k \Big[ W_{m}^2 b_{i_m}^2
  + 2 \sum_{j=0}^{m-1} W_m W_j b_{i_m} b_{i_j} \Big] \Big)-
  x_i^2\:.
  \label{eq:direct_variance_3}
\end{equation}
We can distribute the random walk probability to yield an expanded
form of the variance:
\begin{equation}
  \sigma_i^2 = \sum_{\nu} P_{\nu} \sum_{m=0}^k W_{m}^2 b_{i_m}^2 + 2
  \sum_{\nu} P_{\nu} \sum_{m=0}^k \sum_{j=0}^{m-1} W_m W_j b_{i_m}
  b_{i_j} - x_i^2\:.
  \label{eq:direct_variance_3_2}
\end{equation}
In particular, we will isolate the first term of the variance by
expanding it:
\begin{equation}
  \sum_{\nu} P_{\nu} \sum_{m=0}^k W_{m}^2 b_{i_m}^2 =
  \sum_{k=0}^{\infty}\sum_{i_1}^{N}\sum_{i_2}^{N}\ldots \sum_{i_k}^{N}
  p_{i,i_1}p_{i_1,i_2}\ldots p_{i_{k-1},i_k}
  w^2_{i,i_1}w^2_{i_1,i_2}\ldots w^2_{i_{k-1},i_k} b_{i_k}^2\:.
  \label{eq:direct_variance_4}
\end{equation}
Using this expansion, we can arrive at a more natural
reason for enforcing $\rho(\ve{H}) < 1$ for our Monte Carlo method to
converge. Per the Hadamard product, we can concatenate the summation
in Eq~(\ref{eq:direct_variance_4}):
\begin{equation}
  (\ve{P} \circ \ve{W} \circ \ve{W})^k =
  \sum_{k=0}^{\infty}\sum_{i_1}^{N}\sum_{i_2}^{N}\ldots \sum_{i_k}^{N}
  p_{i,i_1}p_{i_1,i_2}\ldots p_{i_{k-1},i_k}
  w^2_{i,i_1}w^2_{i_1,i_2}\ldots w^2_{i_{k-1},i_k}\:.
  \label{eq:double_weighted_decomposition}
\end{equation}
If we assign $\ve{G} = \ve{P} \circ \ve{W} \circ \ve{W}$ as in
Eq~(\ref{eq:neumann_ulam_decomposition}), we then have:
\begin{equation}
   (\ve{P} \circ \ve{W} \circ \ve{W})^k =
  \sum_{k=0}^{\infty}\sum_{i_1}^{N}\sum_{i_2}^{N}\ldots
  \sum_{i_k}^{N}g_{i,i_1}g_{i_1,i_2}\ldots g_{i_{k-1},i_k}\:,
  \label{eq:direct_variance_6}
\end{equation}
which is the general Neumann series for $\ve{G}$,
\begin{equation}
  \ve{T} = \sum_{k=0}^{\infty} \ve{G}^k\:,
  \label{eq:variance_neumann_series}
\end{equation}
where $\ve{T} = (\ve{I}-\ve{G})^{-1}$. We can then insert $T$ back
into the variance formulation for a more concise definition:
\begin{equation}
  \sigma^2_i = (\ve{T}\ve{b}^2)_i + 2 \sum_{\nu} P_{\nu} \sum_{m=0}^k
  \sum_{j=0}^{m-1} W_m W_j b_{i_m} b_{i_j} - x_i^2\:.
  \label{eq:direct_variance_5}
\end{equation}
We can relate $\ve{G}$ to $\ve{H}$ by noting that $\ve{G}$ simply
contains an additional Hadamard product with the weight matrix. The
Hadamard product has the property that:
\begin{equation}
  ||\ve{H} \circ \ve{W}|| \geq ||\ve{H}||\ ||\ve{W}||\:.
  \label{eq:hadamard_inequality}
\end{equation}
Using the norm property of the Hadamard product and
Eq~(\ref{eq:neumann_ulam_decomposition}), we can define the norm of
$\ve{W}$ as:
\begin{equation}
  \frac{||\ve{H}||}{||\ve{P}||} \geq ||\ve{W}||\:.
  \label{eq:weight_norm}
\end{equation}
Choosing the 2-norm of the operators, if we assume symmetry of
$\ve{H}$ and therefore $\ve{P}$, $\ve{W}$, and $\ve{G}$, we have
\cite{leveque_finite_2007}:
\begin{equation}
  ||\ve{H}||_2 = \rho(\ve{H})\:.
  \label{eq:matrix_2_norm}
\end{equation}
The row normalized probability matrix will yield a norm of 1 giving
the following inequality for relating $\ve{G}$ and $\ve{H}$:
\begin{equation}
  ||\ve{G}||_2 \geq ||\ve{H}||^2_2\:,
  \label{eq:variance_norm_inequality}
\end{equation}
or
\begin{equation}
  \rho(\ve{G}) \geq \rho(\ve{H})^2\:.
  \label{eq:variance_norm_inequality_2}
\end{equation}
Using these relations to analyze Eq~(\ref{eq:direct_variance_5}), we
see that if $\rho(\ve{G}) > 1$, then the first sum in
Eq~(\ref{eq:variance_neumann_series}) will not converge and an
infinite variance will arise as the elements of $\ve{T}$ become
infinite in Eq~(\ref{eq:direct_variance_5}). We must restrict $\ve{G}$
to alleviate this and therefore restrict $\ve{H}$ due to
Eq~(\ref{eq:variance_norm_inequality_2}) with $\rho(\ve{H}) < 1$ so that
our expectation values for the solution may have a finite variance.

\section{Collision Estimator Variance\ }
\label{subsec:forward_variance}
We can compute the variance of the collision estimator in the same way
as the direct estimator for each component in the solution where now:
\begin{equation}
  \sigma_j^2 = \sum_{\nu} P_{\nu} \sum_{m=0}^k W_{m}^2 \delta_{i_m,j}
  + 2 \sum_{\nu} P_{\nu} \sum_{m=0}^k \sum_{l=0}^{m-1} W_m W_l
  \delta_{i_m,j} \delta_{i_l,j} - x_j^2\:.
  \label{eq:collision_variance_1}
\end{equation}
In this case, in the second summation $m \neq l$ for all states due to
the form of the summation and therefore the combination of delta
functions, $\delta_{i_m,j}$ of $\delta_{i_l,j}$, will only be nonzero
for random walks that are in the current solution state (when $i_m =
i_l = j$) and other states visited do not contribute to the variance
of the current state (when $i_m \neq i_l$). This is important as it
shows a decoupling of the variance among the solution vector states,
meaning that the variance in solution state $i$ does not depend on the
variance in solution state $j$.

Expanding the transition probabilities in the first sum again yields a
Neumann series in the same form as that explored for the forward
estimator, this time with the terms in reverse order and the
introduction of the Kronecker delta:
\begin{multline}
  \sigma_j^2 = \sum_{k=0}^{\infty}\sum_{i_1}^{N}\sum_{i_2}^{N}\ldots
  \sum_{i_k}^{N} p_{i_k,i_{k-1}}\ldots p_{i_2,i_1} p_{i_1,i_0}
  w^2_{i_k,i_{k-1}}\ldots w^2_{i_2,i_1} w^2_{i_1,i_0}
  b^2_{i_0}\delta_{i_k,j} + \\ 2 \sum_{\nu} P_{\nu} \sum_{m=0}^k
  \sum_{l=0}^{m-1} W_m W_l \delta_{i_m,j} \delta_{i_l,j} - x_j^2\:.
  \label{eq:collision_variance_2}
\end{multline}
If we again define $\ve{G} = \ve{P} \circ \ve{W} \circ \ve{W}$, we
then have:
\begin{multline}
  \sigma^2_j = \sum_{k=0}^{\infty}\sum_{i_1}^{N}\sum_{i_2}^{N}\ldots
  \sum_{i_k}^{N}g_{i_{k},i_{k-1}} \ldots
  g_{i_2,i_1}g_{i_1,i_0}b_{i_0}\delta_{i_k,j} + \\ 2 \sum_{\nu}
  P_{\nu} \sum_{m=0}^k \sum_{l=0}^{m-1} W_m W_l \delta_{i_m,j}
  \delta_{i_l,j} - x_j^2\:,
\label{eq:collision_variance_3}
\end{multline}
where now $\ve{G}$ in this case is the transpose of that used in
Eq~(\ref{eq:direct_variance_6}). The Kronecker delta again implies
that the variance contribution from each random walk will only be in
its current state and for many random walks, many starting weights of
$b_{i_0}$ will contribute to the variance in the same way as they
contribute to the solution tally. As the transpose is formed and the
Neumann series of $\ve{G}$ is in reverse order in
Eq~(\ref{eq:collision_variance_3}) relative to its formulation in the
forward estimator, then Eq~(\ref{eq:collision_variance_3}) and
Eq~(\ref{eq:direct_variance_6}) are equivalent and therefore the
collision estimator is bound by the same restrictions on
$\rho(\ve{G})$ and $\rho(\ve{H})$ to ensure that the variance is
finite.

\section{Expected Value Estimator Variance\ }
\label{subsubsec:expected_value_estimator_variance}
Following the collision estimator variance, the expected value
estimator variance is given as:
\begin{multline}
  \sigma_j^2 = \sum_{\nu} P_{\nu} b_j^2 + 2 \sum_{\nu} P_{\nu} b_j
  \sum_{m=0}^k W_{m} h_{j,i_m} +\\ \sum_{\nu} P_{\nu} \sum_{m=0}^k
  W_{m}^2 h_{j,i_m}^2 + 2 \sum_{\nu} P_{\nu} \sum_{m=0}^k
  \sum_{l=0}^{m-1} W_m W_l h_{j,i_m} h_{j,i_l} - x_j^2\:.
  \label{eq:expected_value_variance_1}
\end{multline}
As the delta functions are no longer present, the final summation in
Eq~(\ref{eq:expected_value_variance_1}) contains contributions for all
valid state combinations in the same row of the iteration matrix and
therefore the variance of each state in the solution tally is coupled
to a group of other states as defined by the sparsity pattern of the
iteration matrix. This coupling of variances is expected due to the
deterministic averaging used to generate the expected value estimator.