### 2013-08-26 doc/dissertation/document/frontmatter/abstract.tex

Predictive modeling and simulation is an important and developing
component of modern nuclear reactor analysis. New high performance
computing hardware and algorithms that leverage this hardware are
needed to help meet the goals and requirements of engineering
analysis. Current practices in neutron transport and fluid flow
leverage methods based on subspace iterations to achieve good
convergence properties and quality parallel scaling on large
machines. However, the next generation of machines is expected to
possess both high levels of concurrency and low memory availability
per core compared to today's machines with a potentially high rate of
faults. Therefore, we seek new algorithms to solve these problems in
reactor analysis that have the potential to perform well on
forthcoming hardware with the goal of improving both iterative
performance and parallel scalability.

This work researches and develops Monte Carlo Synthetic Acceleration
(MCSA) methods as a new class of solution techniques for discrete
neutron transport and fluid flow problems. Monte Carlo Synthetic
Acceleration methods use a traditional Monte Carlo process to
approximate the solution to the discrete problem as a means of
accelerating traditional fixed-point methods. To apply these methods
to neutronics and fluid flow and determine the feasibility of these
methods on modern hardware, three complementary research and
development exercises are performed.

First, solutions to the $SP_N$ discretization of the linear Boltzmann
neutron transport equation are obtained using MCSA with a difficult
criticality calculation for a light water reactor fuel assembly used
as the driving problem. It was discovered that basic Jacobi-based
preconditioning as used in previous work was not sufficient to achieve
convergence due to the dominating effect of the light water
moderator. To enable MCSA as a solution technique a group of modern
preconditioning strategies are researched. MCSA was verified to
produce the correct solution for the fuel assembly criticality problem
when compared to conventional Krylov methods and demonstrated improved
iterative performance over GMRES by converging in fewer iterations
when using the same preconditioning. The new preconditioning
techniques, however, create unwanted growth in both memory consumption
and computation time for these problems.

Second, solutions to the compressible Navier-Stokes equations were
obtained by developing the Forward-Automated Newton-MCSA (FANM) method
for nonlinear systems based on Newton's method. Three difficult fluid
benchmark problems in both convective and driven flow regimes were
used to drive the research and development of the method. It was found
for all benchmarks that FANM produced identical solutions when
compared to those generated by modern Newton-Krylov methods. For each
benchmark, performance comparisons between FANM and Newton-Krylov were
also conducted. For 8 out of 12 benchmark cases, it was found that
FANM had better iterative performance than the Newton-Krylov method by
converging the nonlinear residual in fewer linear solver iterations
with the same preconditioning.

Third, a new domain decomposed algorithm to parallelize MCSA aimed at
leveraging leadership-class computing facilities was developed by
utilizing parallel strategies from the radiation transport
community. The new algorithm utilizes the Multiple-Set
Overlapping-Domain strategy in an attempt to reduce parallel overhead
and add a natural element of replication to the algorithm. A large
number of parallel scaling studies using a neutron diffusion problem
were performed on the Titan Cray XK7 machine at Oak Ridge National
Laboratory with up to 65,356 cores used in the strong scaling studies
and 131,072 cores used in the weak scaling studies. It was found that
for the current implementation of MCSA, both weak and strong scaling
improved on that observed for production implementations of Krylov
methods. In addition, the replication in the algorithm was able to
boost parallel efficiencies by up to 20\% at large core counts in the
strong scaling studies and enhance time to solution in the weak
scaling stduies by using the principle of superposition. Finally, it
was discovered that MCSA exhibits the most scalable behavior when it
is formulated as a stochastic realization of an additive Schwarz
method.

Each of these research and development activities meet the goals of
this work by producing methods that in many cases were demonstrated to
converge in fewer iterations than conventional methods and exhibit
better parallel efficiencies at high levels of concurrency. All the
components of this work are required to enable future simulation of
nonlinear transport phenomena on the next generation of hardware. To
enable high performance simulations with FANM, we require the MCSA
techniques developed in the research on the neutron transport problem
and in turn, effectively leveraging MCSA on leadership-class hardware
requires the new parallel algorithm developed by this work. With the
further development of methods based on this work, future nuclear
reactor simulations may potentially be enabled that have an improved
time to solution and more efficiently utilize the available
computational resources.